### 2018-06-12 paper.tex
added paper

%%%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[10pt, sigplan]{acmart}
%%\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[sigplan,10pt]{acmart}\settopmatter{}

%% Conference information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
%\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
%\acmYear{2017}
%\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
%\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style

%\bibliographystyle{ACM-Reference-Format}

%% Citation style
%\citestyle{acmauthoryear}  %% For author/year citations
\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format must update the
%% '\documentclass' and topmatter commands above; see
%% 'acmart-pacmpl-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{ifthen}
\usepackage{pgfplots}
\usepackage{listings}
\usepackage{multirow}
\usepgfplotslibrary{statistics}
\usepackage{dblfloatfix} %enable fig at bottom of page
%\input{macros.tex}

\usepackage{xcolor}
\newcommand{\todo}[1]{\color{orange}\fbox{\bfseries\sffamily\scriptsize TODO:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
\newcommand{\sd}[1]{\color{red}\fbox{\bfseries\sffamily\scriptsize Stef:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
\newcommand{\sk}[1]{\color{blue}\fbox{\bfseries\sffamily\scriptsize Sophie:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
\newcommand{\cba}[1]{\color{purple}\fbox{\bfseries\sffamily\scriptsize Clement:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
%\newcommand*{rotatebox{75}}
\input{macros}

\begin{document}

%%%
% Legend definition (so I can change it once for all)
%%%

\def\legBase{Baseline\xspace}
\def\legSmartSyntaxPlug{Syntax\xspace}
\def\legSlangPlug{Plugin\xspace}
\def\legSlang{Slang\xspace}
\def\legRTL{Slang+RTL\xspace}
\def\LLVMMac{LLVM-Mac\xspace}
\def\GCCLinux{GCC-Linux\xspace}
\def\plotHeight{6.8cm}

%%%
% End Legend
%%%

%% Title information
\title[Garbage Collection Evaluation Infrastructure]{Garbage Collection Evaluation Infrastructure for the Cog VM}

%% Author with single affiliation.
\author{Sophie Kaleba}
                                        %% can be repeated if necessary
\affiliation{
  %\position{Position1}
  \department{Master student}              %% \department is recommended
  \institution{Université de Lille 1}            %% \institution is required
 % \streetaddress{Street1 Address1}
  \city{Lille}
  %\state{France}
  %\postcode{Post-Code1}
  \country{France}                    %% \country is recommended
}
\email{sophie.kaleba@etudiant.univ-lille1.fr}          %% \email is recommended

%% Author with two affiliations and emails.
\author{Cl\'ement B\'era}
\affiliation{
  % \position{}
	\department{Software Languages Lab}              %% \department is recommended
	\institution{Vrije Universiteit Brussel}            %% \institution is required
	\city{Brussel}
  % \state{}
  % \postcode{}
	\country{Belgium}                    %% \country is recommended
}
\email{clement.bera@vub.ac.be}          %% \email is recommended

\begin{abstract}

\end{abstract}

\keywords{Garbage Collector, Virtual machine, Managed runtime}  %% \keywords are mandatory in final camera-ready submission

\maketitle

\section{Introduction}
\label{sec:intro}

%% Bibliography
\bibliographystyle{alpha}
\bibliography{sista}

\end{document}

### 2018-06-12 paper.tex
added sec 2-3

@@
 \newcommand{\todo}[1]{\color{orange}\fbox{\bfseries\sffamily\scriptsize TODO:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
-\newcommand{\sd}[1]{\color{red}\fbox{\bfseries\sffamily\scriptsize Stef:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
 \newcommand{\sk}[1]{\color{blue}\fbox{\bfseries\sffamily\scriptsize Sophie:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
@@
 \begin{document}
-
-%%%
-% Legend definition (so I can change it once for all)
-%%%
-
-\def\legBase{Baseline\xspace}
-\def\legSmartSyntaxPlug{Syntax\xspace}
-\def\legSlangPlug{Plugin\xspace}
-\def\legSlang{Slang\xspace}
-\def\legRTL{Slang+RTL\xspace}
-\def\LLVMMac{LLVM-Mac\xspace}
-\def\GCCLinux{GCC-Linux\xspace}
-\def\plotHeight{6.8cm}
-
-%%%
-% End Legend
-%%%
 
@@
 
+Trying to build a more advanced algo, need to measure..
+ref impl to compare
+Try to build benchmarks 
+->
+
+\section{Reference Implementations}
+\label{sec:ref}
+
+To evaluate new full GC algorithms, we need to compare it against existing algorithms.
+Two of the most common GC algorithms are Mark-Compact and Mark-Sweep.
+We describe here the current Mark-Compact implementation, which has been in production for the past few years, and an implementation of Mark-Sweep we introduced as a reference earlier this year.
+
+\subsection{Mark-Compact}
+
+Cog's Mark-Compact algorithm, named \emph{SpurPlanningCompactor}, implements the classic planning compaction algorithm for Spur.
+It uses the fact that there is room for a forwarding pointer in all objects to store the eventual position of an object in the first field.
+It therefore first locates a large free chunk, or eden or a memory segment, to use as the savedFirstFieldsSpace, which it uses to store the first fields of objects that will be compacted.
+It then makes at least three passes through the heap.
+
+The first pass plans where live movable objects will go, copying their first field to the next slot in savedFirstFieldsSpace, and setting their forwarding pointer to point to their eventual location (see planCompactSavingForwarders).
+The second pass updates all pointers in live pointer objects to point to objects' final destinations, including the fields in savedFirstFieldsSpace (see updatePointers and updatePointersInMobileObjects).
+The third pass moves objects to their final positions, unmarking objects, and restoring saved first fields as it does so (see copyAndUnmark: and copyAndUnmarkMobileObjects).
+If the forwarding fields of live objects in the to-be-moved portion of the entire heap won't fit in savedFirstFieldsSpace, then additional passes can be made until the entire heap has been compacted.
+When snapshotting multiple passes are made, but when doing a normal GC only one pass is made.
+
+Each pass uses a three finger algorithm, a simple extension of the classic two finger algorithm with an extra finger used to identify the lowest pinned object between the to and from fingers.
+Objects are moved down, starting at the first free object or chunk, provided that they fit below the lowest pinned object above the to finger.
+When an object won't fit the to finger is moved above the pinned object and the third finger is reset to the next pinned object below the from finger, if any.
+
+\subsection{Mark-Sweep}
+
+Cog's Mark-Compact algorithm, named \emph{SpurSweeper}, is a sweep-only algorithm, setting the compactor to SpurSweeper effectively changes the fullGC to a mark-sweep non-moving algorithm.
+
+SpurSweeper is a reference implementation if one wants to evaluate GC performance and compare it to a Mark-Sweep.
+It's also the only non-moving GC available right now which can be convenient for some experiments.
+One of the main reason why it was implemented is because advanced compaction algorithm includes a sweep phase (See SelectiveCompactor for example) and SpurSweeper allows to debug the sweep phase separately.
+
+\subsection{Snapshot discussion}
+Problem with the snapshot. and non compacting algo. Mention G1 and partial compaction too.
+
+=> HybridCompactor
+
+\section{Building Macro-benchmarks}
+\label{sec:bench}
+
+2 directions. One is to build real benchs from real apps. Same direction as Dacapo... To be ported ?
+
+\subsection{Moose benchmark}
+
+Explain all the phases
+
+> Growing heap:
+[GCMooseBench new benchLoadFromMSE: b wildfly].
+benchLoadFromMSE: fileName
+mooseModel := MooseModel new importFromMSEStream: (StandardFileStream readOnlyFileNamed: fileName)
+
+[(GCMooseBench withMooseModel: model) benchExpandProperties]. "optional, increases drastically the model size"
+benchExpandProperties
+	allProperties :=  mooseModel entityStorage collect: [ :anEntity | 
+    		anEntity mooseInterestingEntity complexPropertyPragmas collect: [ :aPragma |
+        		[ (anEntity mooseInterestingEntity perform: aPragma methodSelector) ] on: Error do: ['error'] ] ].
+
+> Stable heap:
+[(GCMooseBench withMooseModel: model) benchAnalysisPyramid].
+benchAnalysisPyramid
+	mooseModel overviewPyramidMetrics.
+[(GCMooseBench withMooseModel: model) benchAnalysisSymbol].
+benchAnalysisSymbol
+	mooseModel allMethods symbolsUsedInName
+[(GCMooseBench withMooseModel: model) benchAnalysisDeprecation].
+benchAnalysisDeprecation
+	mooseModel allModelClasses select: [ :each |
+    		(each isAnnotatedWith: 'Deprecated') and: [ each clientTypes notEmpty ] ].
+
+> Shrinking heap:
+[(GCMooseBench withMooseModel: model) benchRelease].
+benchRelease
+	mooseModel := nil.
+	allProperties := nil.
+	3 timesRepeat: [Smalltalk garbageCollect].
+
+\subsection{Configurable benchmark}
+
+Sophie's section
+
+\section{Experimental results}
+\label{sec:valid}
+
+It would be nice to evaluate both Sweeper and Planning on what we've just discussed.
+
+\section{Related and Future Work}
+
+\paragraph{GC benchmark suite.}
+Dacapo + see ref in Dacapo.
+
+\paragraph{ACDC.}
+ACDC-js (Sophie you write a paragraph on that.
+
+\paragraph{Future work: more benchmarks.}
+need to add more benchmarks based on Dacapo ?
+
+\paragraph{Future work: stress GC tests.}
+stress cases from Configurable
+
 %% Bibliography

### 2018-06-12 paper.tex
wrote abstract

@@
 
+%% Author with two affiliations and emails.
+\author{Eliot Miranda}
+\affiliation{
+   \position{Virtual Machine Architect}
+	%\department{Virtual Machine architect}              %% \department is recommended
+	\institution{Feenk}            %% \institution is required
+	\city{San Francisco}
+  % \state{}
+  % \postcode{}
+	\country{California}                    %% \country is recommended
+}
+\email{eliot.miranda@gmail.com}          %% \email is recommended
+
 \begin{abstract}
-
+One of the next step to improve the Cog virtual machine, the default virtual machine for multiple programming languages such Pharo, Squeak and Newspeak, is to decrease the garbage collection pause time.
+A benchmarking infrastructure and reference garbage collection algorithm implementations are required to evaluate the performance of a new algorithm and compare it.
+Cog features a Mark-Compact algorithm, used in production, and we introduced a Mark-Sweep algorithm a the two reference algorithms.
+Benchmarks are built using two different approaches.
+Firstly, we turned code from memory intensive deployed applications into benchmarks to simulate real-world applications.
+Secondly, we built a configurable benchmark which simulates an application with different heap properties to be able to stress specific aspects of the memory management.
+We then evaluated the two reference algorithms on the infrastructure built to have reference benchmark results.
 \end{abstract}
 
-\keywords{Garbage Collector, Virtual machine, Managed runtime}  %% \keywords are mandatory in final camera-ready submission
+\keywords{Benchmark, Garbage Collector, Virtual machine, Managed runtime}  %% \keywords are mandatory in final camera-ready submission
 
@@
 
-Trying to build a more advanced algo, need to measure..
-ref impl to compare
-Try to build benchmarks 
-->
+- Trying to build a more advanced full GC algo, need to measure..
+- ref impl to compare
+- Try to build benchmarks, both from real app and to stress specific aspects.
+- We end by applying the bench on reference impl to validate the behavior is as expected.
 
@@
 
-=> HybridCompactor
+=> HybridCompactor solution.
 

### 2018-06-12 paper.tex
intro beginning

@@
 
-- Trying to build a more advanced full GC algo, need to measure..
-- ref impl to compare
-- Try to build benchmarks, both from real app and to stress specific aspects.
-- We end by applying the bench on reference impl to validate the behavior is as expected.
+The Cog virtual machine (VM), the default VM for multiple programming languages such Pharo, Squeak and Newspeak, currently features in production a stop-the-world Mark-Compact algorithm as the full garbage collector (GC) algorithm.
+The algorithm has a high throughput but a high pause time during which the application is not responsive (the GC interrupts the application for multiple seconds on modern Macbooks starting from multiple Gbs heaps).
+For interactive applications, a new algorithm is required with a smaller pause time.
 
-\section{Reference Implementations}
-\label{sec:ref}
+Building and tuning a new GC algorithm requires to evaluate its behavior. A benchmarking infrastructure is required to do so. To build GC benchmarks, we took two approaches.
 
-To evaluate new full GC algorithms, we need to compare it against existing algorithms.
-Two of the most common GC algorithms are Mark-Compact and Mark-Sweep.
-We describe here the current Mark-Compact implementation, which has been in production for the past few years, and an implementation of Mark-Sweep we introduced as a reference earlier this year.
+Firstly, we contacted multiple companies using the Cog VM in production on memory intensive application (>1Gb heaps) and built benchmarks out of their deployed application.
+As an example, we will discuss in the paper the Moose benchmarks implemented in collaboration with Feenk\footnote{https://feenk.com/}.
+Part of the business of Feenk consists in analysing software written in multiple programming languages using the open-source framework Moose~\cite{MoosePaper1,MooseBook1}.
+Using Moose, the application parses the software to analyse into a model, performs analysis on it and lastly release the model.
+This behavior was turned into three benchmarks, growing, accessing and shrinking the heap.
 
-\subsection{Mark-Compact}
+Secondly, \todo{Sophie :-) a few sentence on your bench}
 
-Cog's Mark-Compact algorithm, named \emph{SpurPlanningCompactor}, implements the classic planning compaction algorithm for Spur.
-It uses the fact that there is room for a forwarding pointer in all objects to store the eventual position of an object in the first field.
-It therefore first locates a large free chunk, or eden or a memory segment, to use as the savedFirstFieldsSpace, which it uses to store the first fields of objects that will be compacted.
-It then makes at least three passes through the heap.
+- Gc algo ref
 
-The first pass plans where live movable objects will go, copying their first field to the next slot in savedFirstFieldsSpace, and setting their forwarding pointer to point to their eventual location (see planCompactSavingForwarders).
-The second pass updates all pointers in live pointer objects to point to objects' final destinations, including the fields in savedFirstFieldsSpace (see updatePointers and updatePointersInMobileObjects).
-The third pass moves objects to their final positions, unmarking objects, and restoring saved first fields as it does so (see copyAndUnmark: and copyAndUnmarkMobileObjects).
-If the forwarding fields of live objects in the to-be-moved portion of the entire heap won't fit in savedFirstFieldsSpace, then additional passes can be made until the entire heap has been compacted.
-When snapshotting multiple passes are made, but when doing a normal GC only one pass is made.
+- We then evaluated the two reference algorithms on the infrastructure built to have reference benchmark results.
 
-Each pass uses a three finger algorithm, a simple extension of the classic two finger algorithm with an extra finger used to identify the lowest pinned object between the to and from fingers.
-Objects are moved down, starting at the first free object or chunk, provided that they fit below the lowest pinned object above the to finger.
-When an object won't fit the to finger is moved above the pinned object and the third finger is reset to the next pinned object below the from finger, if any.
-
-\subsection{Mark-Sweep}
-
-Cog's Mark-Compact algorithm, named \emph{SpurSweeper}, is a sweep-only algorithm, setting the compactor to SpurSweeper effectively changes the fullGC to a mark-sweep non-moving algorithm.
-
-SpurSweeper is a reference implementation if one wants to evaluate GC performance and compare it to a Mark-Sweep.
-It's also the only non-moving GC available right now which can be convenient for some experiments.
-One of the main reason why it was implemented is because advanced compaction algorithm includes a sweep phase (See SelectiveCompactor for example) and SpurSweeper allows to debug the sweep phase separately.
-
-\subsection{Snapshot discussion}
-Problem with the snapshot. and non compacting algo. Mention G1 and partial compaction too.
-
-=> HybridCompactor solution.
+- related and futur
 
@@
 
-\section{Experimental results}
+\section{Reference Implementations}
+\label{sec:ref}
+
+To evaluate new full GC algorithms, we need to compare it against existing algorithms.
+Two of the most common GC algorithms are Mark-Compact and Mark-Sweep.
+We describe here the current Mark-Compact implementation, which has been in production for the past few years, and an implementation of Mark-Sweep we introduced as a reference earlier this year.
+
+\subsection{Mark-Compact}
+
+Cog's Mark-Compact algorithm, named \emph{SpurPlanningCompactor}, implements the classic planning compaction algorithm for Spur.
+It uses the fact that there is room for a forwarding pointer in all objects to store the eventual position of an object in the first field.
+It therefore first locates a large free chunk, or eden or a memory segment, to use as the savedFirstFieldsSpace, which it uses to store the first fields of objects that will be compacted.
+It then makes at least three passes through the heap.
+
+The first pass plans where live movable objects will go, copying their first field to the next slot in savedFirstFieldsSpace, and setting their forwarding pointer to point to their eventual location (see planCompactSavingForwarders).
+The second pass updates all pointers in live pointer objects to point to objects' final destinations, including the fields in savedFirstFieldsSpace (see updatePointers and updatePointersInMobileObjects).
+The third pass moves objects to their final positions, unmarking objects, and restoring saved first fields as it does so (see copyAndUnmark: and copyAndUnmarkMobileObjects).
+If the forwarding fields of live objects in the to-be-moved portion of the entire heap won't fit in savedFirstFieldsSpace, then additional passes can be made until the entire heap has been compacted.
+When snapshotting multiple passes are made, but when doing a normal GC only one pass is made.
+
+Each pass uses a three finger algorithm, a simple extension of the classic two finger algorithm with an extra finger used to identify the lowest pinned object between the to and from fingers.
+Objects are moved down, starting at the first free object or chunk, provided that they fit below the lowest pinned object above the to finger.
+When an object won't fit the to finger is moved above the pinned object and the third finger is reset to the next pinned object below the from finger, if any.
+
+\subsection{Mark-Sweep}
+
+Cog's Mark-Compact algorithm, named \emph{SpurSweeper}, is a sweep-only algorithm, setting the compactor to SpurSweeper effectively changes the fullGC to a mark-sweep non-moving algorithm.
+
+SpurSweeper is a reference implementation if one wants to evaluate GC performance and compare it to a Mark-Sweep.
+It's also the only non-moving GC available right now which can be convenient for some experiments.
+One of the main reason why it was implemented is because advanced compaction algorithm includes a sweep phase (See SelectiveCompactor for example) and SpurSweeper allows to debug the sweep phase separately.
+
+\subsection{Snapshot discussion}
+Problem with the snapshot. and non compacting algo. Mention G1 and partial compaction too.
+
+=> HybridCompactor solution.
+
+\section{Reference Results}
 \label{sec:valid}

### 2018-06-12 paper.tex
well, thats an intro

@@
 Firstly, we contacted multiple companies using the Cog VM in production on memory intensive application (>1Gb heaps) and built benchmarks out of their deployed application.
-As an example, we will discuss in the paper the Moose benchmarks implemented in collaboration with Feenk\footnote{https://feenk.com/}.
+As an example, we will discuss in the paper in Section \ref{sec:mooseBench} the Moose benchmarks implemented in collaboration with Feenk\footnote{https://feenk.com/}.
 Part of the business of Feenk consists in analysing software written in multiple programming languages using the open-source framework Moose~\cite{MoosePaper1,MooseBook1}.
@@
 
-Secondly, \todo{Sophie :-) a few sentence on your bench}
+Secondly, \todo{Sophie :-) a few sentence on your bench + ref} \ref{sec:confBench}
 
-- Gc algo ref
+To evaluate a new algorithm, we also need to compare it to reference implementations.
+The Cog VM features a Mark-Compact algorithm in production.
+The compaction phase is quite specific since it moves objects in memory and spend time updating references.
+In addition to this algorithm, we implemented a Mark-Sweep, which does not move objects in memory and is quicker to perform.
+Section \ref{sec:ref} details the two implementations.
 
-- We then evaluated the two reference algorithms on the infrastructure built to have reference benchmark results.
+We evaluated both algorithm on the benchmarks we built to provide results of the reference algorithms on the infrastructure built.
 
-- related and futur
+We conclude the paper by discussing some related work, the GC benchmarks suites for Java and the ACDC benchmarks, and some future work.
 
@@
 \subsection{Moose benchmark}
+\label{sec:mooseBench}
 
@@
 \subsection{Configurable benchmark}
+\label{sec:confBench}
 

### 2018-06-13 paper.tex
added todo for the other author
added ref and section bench

@@
 
-The Cog virtual machine (VM), the default VM for multiple programming languages such Pharo, Squeak and Newspeak, currently features in production a stop-the-world Mark-Compact algorithm as the full garbage collector (GC) algorithm.
+The Cog virtual machine (VM), the default VM for multiple programming languages such Pharo \cite{PharoByExample}, Squeak \cite{SqueakByExample} and Newspeak \cite{NewspeakOopsla}, currently features in production a stop-the-world Mark-Compact algorithm as the full garbage collector (GC) algorithm.
 The algorithm has a high throughput but a high pause time during which the application is not responsive (the GC interrupts the application for multiple seconds on modern Macbooks starting from multiple Gbs heaps).
@@
 
-\section{Building Macro-benchmarks}
+\section{Building Benchmarks}
 \label{sec:bench}
 
-2 directions. One is to build real benchs from real apps. Same direction as Dacapo... To be ported ?
+To build GC benchmarks, we took two approaches.
+We asked companies for their deployed memory intensive applications and turned them into benchmarks.
+Multiple benchmarks in this category include closed-source code.
+In Section \ref{sec:mooseBench}, we describe how we turned a specific application into a benchmark with the help of the company (this specific benchmark does not include closed-source code).
+Then, to stress the GC on specific aspects, we built a configurable benchmark that based on the configuration emulates heaps with different properties.
 
@@
 
-Explain all the phases
+Part of the business of Feenk consists in analyzing softwares.
+To do so, the application parses a mse file into a model, performs some analysis, and then releases the model and generates analysis results.
+For business, the application analyzes closed-sourced application of companies.
+To build an open-source benchmark, we analyse open-source applications instead of closed-source ones.
 
-> Growing heap:
-[GCMooseBench new benchLoadFromMSE: b wildfly].
-benchLoadFromMSE: fileName
+We built six different benchmarks that falls into three categories: growing, stable and shrinking heaps.
+The benchmarks were built on top of the stable Moose image\footnote{http://www.moosetechnology.org/}.
+
+\paragraph{Growing heap.} The growing heap benchmarks increase the heap size as they are performed, mainly creating objects. We built two benchmarks:
+\begin{itemize}
+	\item \emph{LoadFromMSE}: parses a mse file into a moose model, effectively loading a graph of object. This is the only benchmark not taking a moose model as a parameter.
+	\item \emph{ExpandProperties}: computes all interesting properties of the model and store them, effectively loading a graph of object connected in many points to the original graph. 
+\end{itemize}
+
+\begin{code}
+GCMooseBench >> benchLoadFromMSE: fileName
 	mooseModel := MooseModel new importFromMSEStream: (StandardFileStream readOnlyFileNamed: fileName)
 
-[(GCMooseBench withMooseModel: model) benchExpandProperties]. "optional, increases drastically the model size"
-benchExpandProperties
+GCMooseBench >>benchExpandProperties
 	allProperties :=  mooseModel entityStorage collect: [ :anEntity | 
@@
         		[ (anEntity mooseInterestingEntity perform: aPragma methodSelector) ] on: Error do: ['error'] ] ].
+\end{code}
 
-> Stable heap:
-[(GCMooseBench withMooseModel: model) benchAnalysisPyramid].
+\paragraph{Stable heap.} The stable heap benchmarks perform random memory accesses in a heap which size remains approximately the same. We built three benchmarks:
+
+\begin{itemize}
+	\item \emph{Pyramid}: computes an overview of the pyramid metrics of the software.
+	\item \emph{Symbol}: computes all the symbols used in name. \todo{check}
+	\item \emph{Deprecation}: computes all the deprecated code.
+\end{itemize}
+
+\begin{code}
 benchAnalysisPyramid
 	mooseModel overviewPyramidMetrics.
-[(GCMooseBench withMooseModel: model) benchAnalysisSymbol].
+
 benchAnalysisSymbol
 	mooseModel allMethods symbolsUsedInName
-[(GCMooseBench withMooseModel: model) benchAnalysisDeprecation].
+
 benchAnalysisDeprecation
@@
     		(each isAnnotatedWith: 'Deprecated') and: [ each clientTypes notEmpty ] ].
+\end{code}
 
-> Shrinking heap:
-[(GCMooseBench withMooseModel: model) benchRelease].
+\paragraph{Shrinking heap.} The shrinking heap benchmark frees part of the heap and performs a GC.
+We built only one benchmark, \emph{Release}, which removes references to the graph of objects and performs three garbage collections.
+
+\begin{code}
 benchRelease
@@
 	3 timesRepeat: [Smalltalk garbageCollect].
+\end{code}
 
@@
 
+\todo{Sophie}
 Sophie's section
@@
 Cog's Mark-Compact algorithm, named \emph{SpurSweeper}, is a sweep-only algorithm, setting the compactor to SpurSweeper effectively changes the fullGC to a mark-sweep non-moving algorithm.
-
-SpurSweeper is a reference implementation if one wants to evaluate GC performance and compare it to a Mark-Sweep.
-It's also the only non-moving GC available right now which can be convenient for some experiments.
-One of the main reason why it was implemented is because advanced compaction algorithm includes a sweep phase (See SelectiveCompactor for example) and SpurSweeper allows to debug the sweep phase separately.
+It iterates over all old space in linear order.
+Each time an unmarked object or a free chunk is met, it coalesces it with subsequent ones and updates the free lists with a new larger free chunk.
 
 \subsection{Snapshot discussion}
-Problem with the snapshot. and non compacting algo. Mention G1 and partial compaction too.
 
-=> HybridCompactor solution.
+On top of the Cog VM, snapshots can be performed to persist the given state of the heap.
+Since snapshots are persisted, they should use a low amount of memory to avoid using too many bytes on disk.
+Non compacting algorithms, such as the Mark-Sweep, or algorithms compacting only part of the heap, such as garbage first~\cite{G1}, are a problem in this context.
+The heap, when collected by the Mark-Sweep, has a lot of free chunks when the heap size varies a lot leading to a large snapshot.
+However, non compacting have also interesting advantages (lower garbage collection pause time for example).
+
+To be able to have a garbage collector with a low pause time and still be able to perform efficiently snapshots, we design a hybrid compactor solution.
+The VM tells the compactor if the garbage collection is performed for snapshot or not, and the compactor chooses to apply one algorithm or the other to have low compaction pause time or to snapshot with a low memory footprint.
 
@@
 \paragraph{GC benchmark suite.}
-Dacapo + see ref in Dacapo.
+Dacapo + see ref in Dacapo. \cite{DacapoBench}
+\todo{Sophie}
 
 \paragraph{ACDC.}
-ACDC-js (Sophie you write a paragraph on that.
+ACDC-js (Sophie you write a paragraph on that. \cite{ACDCJS}
 

### 2018-06-13 paper.tex
c

@@
 \begin{code}
-benchAnalysisPyramid
+GCMooseBench >> benchAnalysisPyramid
 	mooseModel overviewPyramidMetrics.
 
-benchAnalysisSymbol
+GCMooseBench >> benchAnalysisSymbol
 	mooseModel allMethods symbolsUsedInName
 
-benchAnalysisDeprecation
+GCMooseBench >> benchAnalysisDeprecation
 	mooseModel allModelClasses select: [ :each |
@@
 \begin{code}
-benchRelease
+GCMooseBench >> benchRelease
 	mooseModel := nil.
@@
 
+I think we show exec time, full gc time, scavenge time, compaction time.
+
 \section{Related and Future Work}

### 2018-06-14 paper.tex
add description in intro. fix some typos at the beginning

@@
   \department{Master student}              %% \department is recommended
-  \institution{Université de Lille 1}            %% \institution is required
+  \institution{Universit\'e de Lille 1}            %% \institution is required
  % \streetaddress{Street1 Address1}
@@
 A benchmarking infrastructure and reference garbage collection algorithm implementations are required to evaluate the performance of a new algorithm and compare it.
-Cog features a Mark-Compact algorithm, used in production, and we introduced a Mark-Sweep algorithm a the two reference algorithms.
+Cog features a Mark-Compact algorithm, used in production, and we introduced a Mark-Sweep algorithm a \sk{a word or something is missing here} the two reference algorithms.
 Benchmarks are built using two different approaches.
@@
 Part of the business of Feenk consists in analysing software written in multiple programming languages using the open-source framework Moose~\cite{MoosePaper1,MooseBook1}.
-Using Moose, the application parses the software to analyse into a model, performs analysis on it and lastly release the model.
+Using Moose, the application parses the software to analyse into a model, performs analysis on it and lastly releases the model.
 This behavior was turned into three benchmarks, growing, accessing and shrinking the heap.
 
-Secondly, \todo{Sophie :-) a few sentence on your bench + ref} \ref{sec:confBench}
+Secondly, we implemented a configurable benchmark: the idea is to set-up specific heap structures that will stress the garbage collector.
+A set of specific options is available to the user to tune the type of allocated objects and some of their features.
+These options can be associated with different heap states: growing, stable or shrinking, like in our other approach.
+\todo{fin a ref}
+This implementation and the use-cases are detailed in Section \ref{sec:confBench}
 
@@
 \todo{Sophie}
-Sophie's section
+describe again what is this bench about
+
+\begin{itemize}
+\item Structure of the heap (allocated objects, their size, their liveness)
+\item State of the heap (growing, stable, shrinking)
+\end{itemize}
 
@@
 
+\paragraph{Future work: more options.}
+\sk{maybe? I will add options in the configurator soon. Should I mention it here?}
+
 %% Bibliography

### 2018-06-15 paper.tex
fix detail

@@
 
-We built six different benchmarks that falls into three categories: growing, stable and shrinking heaps.
+We built four different benchmarks that falls into three categories: growing, stable and shrinking heaps.
 The benchmarks were built on top of the stable Moose image\footnote{http://www.moosetechnology.org/}.
@@
 	\item \emph{LoadFromMSE}: parses a mse file into a moose model, effectively loading a graph of object. This is the only benchmark not taking a moose model as a parameter.
-	\item \emph{ExpandProperties}: computes all interesting properties of the model and store them, effectively loading a graph of object connected in many points to the original graph. 
+	\item \emph{ExpandProperties}: computes all interesting properties of the model, discard properties rarely used or consuming too much memory, cache common properties not using too much memory.
+It effectively loads a graph of object connected in many points to the original graph.
 \end{itemize}
 
-\begin{code}
-GCMooseBench >> benchLoadFromMSE: fileName
-	mooseModel := MooseModel new importFromMSEStream: (StandardFileStream readOnlyFileNamed: fileName)
-
-GCMooseBench >>benchExpandProperties
-	allProperties :=  mooseModel entityStorage collect: [ :anEntity | 
-    		anEntity mooseInterestingEntity complexPropertyPragmas collect: [ :aPragma |
-        		[ (anEntity mooseInterestingEntity perform: aPragma methodSelector) ] on: Error do: ['error'] ] ].
-\end{code}
-
-\paragraph{Stable heap.} The stable heap benchmarks perform random memory accesses in a heap which size remains approximately the same. We built three benchmarks:
+\paragraph{Stable heap.} The stable heap benchmarks perform random memory accesses in a heap which size remains approximately the same. We built one benchmark:
 
 \begin{itemize}
-	\item \emph{Pyramid}: computes an overview of the pyramid metrics of the software.
-	\item \emph{Symbol}: computes all the symbols used in name. \todo{check}
-	\item \emph{Deprecation}: computes all the deprecated code.
+	\item \emph{ExpandPropertiesWithCache}: computes all interesting properties of the model, common properties with low memory footprint are already cached and other properties are discarded once computed.
 \end{itemize}
-
-\begin{code}
-GCMooseBench >> benchAnalysisPyramid
-	mooseModel overviewPyramidMetrics.
-
-GCMooseBench >> benchAnalysisSymbol
-	mooseModel allMethods symbolsUsedInName
-
-GCMooseBench >> benchAnalysisDeprecation
-	mooseModel allModelClasses select: [ :each |
-    		(each isAnnotatedWith: 'Deprecated') and: [ each clientTypes notEmpty ] ].
-\end{code}
 
@@
 We built only one benchmark, \emph{Release}, which removes references to the graph of objects and performs three garbage collections.
-
-\begin{code}
-GCMooseBench >> benchRelease
-	mooseModel := nil.
-	allProperties := nil.
-	3 timesRepeat: [Smalltalk garbageCollect].
-\end{code}
 
@@
 
+We evaluated both reference algorithms on the benchmarks built.
+
+\subsection{Moose Benchmarks}
+
+\begin{figure*}[thb]
+	\centering
+    	\begin{subfigure}[b]{.48\textwidth}
+	\includegraphics[width=\linewidth]{figures/load} 
+	\subcaption{Release\vspace{0.5cm}}
+   	\end{subfigure} \hspace{0.03\textwidth} \begin{subfigure}[b]{.48\textwidth}
+	\includegraphics[width=\linewidth]{figures/prop} 
+	\subcaption{ExpandProperties\vspace{0.5cm}}
+   	\end{subfigure}
+
+	\begin{subfigure}[b]{.48\textwidth}
+	\includegraphics[width=\linewidth]{figures/propCache} 
+	\subcaption{ExpandPropertiesWithCache}
+   	\end{subfigure} \hspace{0.03\textwidth}  \begin{subfigure}[b]{.48\textwidth}
+	\includegraphics[width=\linewidth]{figures/release} 
+	\subcaption{release}
+   	\end{subfigure}
+\caption{Moose benchmark results piled up.}
+\label{MooseRes1}
+\end{figure*}
+
+\begin{figure*}[thb]
+\begin{tabular}{|c|c|c|c|c|c|c|c|}
+   \hline
+  & & Total   & Scavenge  & Total Full  & Compaction  & Initial  & Final \\
+ Benchmark & &  Exec. time &  time & GC time &  time &  Heap size & Heap size \\
+  & &  (ms) &  (ms) & (ms) &  (ms) &  (Mb) &  (Mb) \\
+   \hline
+   \multirow{2}{*}{Load} & Sweep 	& 159821 $\pm$ 2218 &	73599 $\pm$ 1281 	& 10863 $\pm$ 726 	&1398 $\pm$ 091 	& 193 $\pm$ 0 	& 959 $\pm$ 9.69 \\
+    				    & Compact 	& 162305 $\pm$ 2790 &	72150 $\pm$ 0782 	& 14091 $\pm$ 972 	&4661 $\pm$ 289 	& 193 $\pm$ 0 	& 909 $\pm$ 9.69 \\
+   \hline
+   \multirow{2}{*}{ExpandProperties} 	& Sweep 		& 232940 $\pm$ 01287 &	55497 $\pm$ 0238 	& 12240 $\pm$ 285 	&1178 $\pm$ 6.66 	& 959 $\pm$ 9.69 & 1888 $\pm$ 0 \\
+    				    			& Compact 	& 183314 $\pm$ 13090 &	53165 $\pm$ 2343 	& 12496 $\pm$ 197 	&3775 $\pm$ 0108 	& 909 $\pm$ 9.69 & 1938 $\pm$ 0 \\
+   \hline
+   ExpandProperties 		& Sweep 		& 510013 $\pm$ 05812 &	64234 $\pm$ 0823 	& 23225 $\pm$ 100 	&2011 $\pm$ 10 	& 1888 $\pm$ 0 	& 1888 $\pm$ 0 \\
+   WithCache  			& Compact 	& 408169 $\pm$ 20444 &	64492 $\pm$ 1234 	& 0 $\pm$ 0 		&0 $\pm$ 0 		& 1938 $\pm$ 0 		& 1938 $\pm$ 0 \\
+   \hline
+   \multirow{2}{*}{Release} 	& Sweep 		& 0606 $\pm$ 16 		&	0 $\pm$ 0 	& 0605 $\pm$ 16 	&316 $\pm$ 12 		& 1888 $\pm$ 0 	& 534 $\pm$ 58 \\
+    				    		& Compact 	& 1092 $\pm$ 02 	&	0 $\pm$ 0 	& 1092 $\pm$ 01 		&806 $\pm$ 03 			& 0193 $\pm$ 0 		& 193 $\pm$ 00 \\
+   \hline
+\end{tabular} 
+\caption{Moose benchmark results with standard errors.}
+\label{MooseRes2}
+\end{figure*}
+
+We evaluated the Moose benchmarks using the reference Mark-Compact and Mark-Sweep algorithm.
+The results are shown in Figure \ref{MooseRes1} and \ref{MooseRes2}.
+In the results, the compaction time includes the time spent in the compact phase in the Mark-Compact and in the sweep phase in the Mark-Sweep.
+Total full GC time is the total time spent in full GCs, including the compaction time.
+Scavenge time is the time spent in scavenges, \emph{i.e.,} garbage collection of young objects only.
+Total execution time is the total execution time, including scavenge and full GC time.
+
+\paragraph{Load.} The Load benchmark grows the memory by 750 Mb with Sweep and by 700 Mb with Compact.
+Compaction time is smaller for the sweep algorithm, making the benchmark a little bit faster in the Mark-Sweep, which is to be expected since Sweep does not need to waste time updating pointers because objects are not moved in memory.
+
+\paragraph{ExpandProperties.} The ExpandProperties benchmark grows the memory by around 1 Gb.
+Compaction time is a little bit better with the sweep algorithm as for the previous benchmark, but execution time is much worse.
+We believe this is due to worse locality of objects.
+
+\paragraph{ExpandPropertiesWithCache.} The ExpandPropertiesWithCache benchmark does not increase the overall memory size.
+Since the compacting GC has compacted the heap before starting the benchmark, there is enough room to allocate the large objects (large objects are allocated directly in old space) required by the benchmark and there is no need for a full GC during the benchmark.
+However, with the sweep algorithm, there are multiple small available memory chunks which are not big enough to hold all the requested large objects, so a full GC is required.
+In addition, execution time is worse, as for the previous benchmark.
+
+\paragraph{Release.} The Release benchmark releases all the graphs.
+With the compacting GC, memory goes back to its original size.
+With Sweep, multiple memory segments are left with a few live objects, forbidding the OS to reclaim those segments.
+Hence, the memory decreases drastically, but not to the original size.
+Compaction time is much faster with Sweep, but as we have discussed, for less efficiency.
+
+\subsection{Sophie section}
+
 It would be nice to evaluate both Sweeper and Planning on what we've just discussed.

### 2018-06-15 paper.tex
minor

@@
 %% Title information
-\title[Garbage Collection Evaluation Infrastructure]{Garbage Collection Evaluation Infrastructure for the Cog VM}
+\title[Garbage Collection Evaluation Infrastructure]{Garbage Collection Evaluation Infrastructure\\ for the Cog VM}
 
@@
 
-\paragraph{Stable heap.} The stable heap benchmarks perform random memory accesses in a heap which size remains approximately the same. We built one benchmark:
-
-\begin{itemize}
-	\item \emph{ExpandPropertiesWithCache}: computes all interesting properties of the model, common properties with low memory footprint are already cached and other properties are discarded once computed.
-\end{itemize}
+\paragraph{Stable heap.} The stable heap benchmarks perform random memory accesses in a heap which size remains approximately the same.
+We built one benchmark, \emph{ExpandPropertiesWithCache}, which computes all interesting properties of the model, common properties with low memory footprint are already cached and other properties are discarded once computed.
 
@@
 
+\todo{Sophie}
 It would be nice to evaluate both Sweeper and Planning on what we've just discussed.

### 2018-06-15 paper.tex
typos

@@
 \begin{abstract}
-One of the next step to improve the Cog virtual machine, the default virtual machine for multiple programming languages such Pharo, Squeak and Newspeak, is to decrease the garbage collection pause time.
+One of the next step to improve the Cog virtual machine, the default virtual machine for multiple programming languages such as Pharo, Squeak and Newspeak, is to decrease the garbage collection pause time.
 A benchmarking infrastructure and reference garbage collection algorithm implementations are required to evaluate the performance of a new algorithm and compare it.
-Cog features a Mark-Compact algorithm, used in production, and we introduced a Mark-Sweep algorithm a \sk{a word or something is missing here} the two reference algorithms.
+Cog features a Mark-Compact algorithm, used in production, and we introduced a Mark-Sweep algorithm as the two reference algorithms.
 Benchmarks are built using two different approaches.
@@
 	\includegraphics[width=\linewidth]{figures/load} 
-	\subcaption{Release\vspace{0.5cm}}
+	\subcaption{Load\vspace{0.5cm}}
    	\end{subfigure} \hspace{0.03\textwidth} \begin{subfigure}[b]{.48\textwidth}
@@
 
-We evaluated the Moose benchmarks using the reference Mark-Compact and Mark-Sweep algorithm.
+We evaluated the Moose benchmarks using the reference Mark-Compact and Mark-Sweep algorithm and by analysing the wildfly code base\footnote{http://www.wildfly.org/}.
 The results are shown in Figure \ref{MooseRes1} and \ref{MooseRes2}.

### 2018-06-15 paper.tex
adds the bit about ACDC

@@
 
-\paragraph{ACDC.}
-ACDC-js (Sophie you write a paragraph on that. \cite{ACDCJS}
+\paragraph{ACDC-JS.}
+ACDC-JS \cite{ACDCJS} is a configurable benchmarking tool designed to specifically measure and stress allocation and deallocation time in the JavaScript memory management model. 
+It builds a heap model based on the behavior of real-time applications and aims at providing significant insights on the JavaScript memory management that cannnot be obtained from standard benchmarking suites.
+The value of specific heap options can be set by the user.
+Their study shows the negative impact of object liveness (time between allocation and last access, leading to the objects to be in different spaces in the heap) and deallocation delay on allocation time.
+It also stresses the trade-offs between memory consumption and allocation latency.
 

### 2018-06-15 paper.tex
fix typo
put figures side by side
describe configurable benchmark

@@
 \todo{fin a ref}
+\sk{add the case we benchmarked with this approach}
 This implementation and the use-cases are detailed in Section \ref{sec:confBench}
@@
 \label{sec:bench}
+\sk{isn't it weird to introduce the built benchmarks without explaining *why* they were chosen?
+Like, why growing, stable, shrinking is significant regarding garbage collection?
+same for my 3 benchs, why would I chose those 3 bench strategies?}
 
@@
 
-\todo{Sophie}
-describe again what is this bench about
+Running a new memory management strategy on real-world applications can give a first insight about its performance.
+Being able to configure a specific set-up, personalized environment, can allow to locally \sk{rephrase} test specific behaviors of the new implementation.
+This configurable benchmarking tool has been implemented to emulate specific heap structures and states to potentially trigger edge cases: the user has available a set of twelve options which can be sorted in two categories:
+
+\paragraph{Heap structure.} A first set of options allows the user to modify the heap structure, i.e.\ the amount and sizes of objects to allocate and the possible interactions with these objects.
+\todo{maybe structure is not actually a good term}
 
 \begin{itemize}
-\item Structure of the heap (allocated objects, their size, their liveness)
-\item State of the heap (growing, stable, shrinking)
+\item \emph{wantedAllocatedSize:} the cumulated size (in Kb) of objects that will be allocated
+\item \emph{minSize and maxSize:} the minimum and maximum size of objects
+\item \emph{dataObjects, weakStructures, classes, compiledMethods:} the ratio for each of these object types.
+\item \emph{accessObjects:} if set to True, access half of the objects allocated on the heap 
+\item \emph{readOnly:} if set to False, modify half of the objects allocated on the heap upon accessing  
+\end{itemize}
+
+\paragraph{Heap state.} The heap state can also be tuned.
+Once objects have been allocated on the heap according to the selected options, it is possible whether to allocate more objects (growing heap), to release objects that have been previously allocated (shrinking heap), or to stay in the current state and potentially accessing and modifying objects that have been previously allocated (stable heap).
+These options echoes the 3 benchmarks depicted in Section \ref{sec:mooseBench}\\
+
+We build three benchmarks out of the potential configurations:
+\begin{itemize}
+\item \emph{FragmentedHeap:} First, objects are allocated one the heap, and some of them are released, leading to a potential memory fragmentation.
+Then, new objets with a minimum size greater than the new free chunks sizes are allocated.
+\item \emph{LargeObjects:} Only large objets with a size greater than 0xFFFF are allocated on the heap.
+\item \emph{Data0bjects:} The majority of objects that will be allocated on the heap will be Data objects, which only contains integer of a specific size (in this benchmark, one byte)
 \end{itemize}
@@
 
+%
 \begin{figure*}[thb]
 	\centering
+
     \begin{subfigure}[b]{.48\textwidth}
 	\includegraphics[width=\linewidth]{figures/load} 
-	\subcaption{Load\vspace{0.5cm}}
-   	\end{subfigure} \hspace{0.03\textwidth} \begin{subfigure}[b]{.48\textwidth}
+	\caption{Load\vspace{0.5cm}}
+   	\end{subfigure}\hspace{0.03\textwidth}% 
+   	\begin{subfigure}[b]{.48\textwidth}
 	\includegraphics[width=\linewidth]{figures/prop} 
-	\subcaption{ExpandProperties\vspace{0.5cm}}
+	\caption{ExpandProperties\vspace{0.5cm}}
+   	\end{subfigure}	
+	\begin{subfigure}[b]{.48\textwidth}
+	\includegraphics[width=\linewidth]{figures/propCache} 
+	\caption{ExpandPropertiesWithCache}
+	\end{subfigure\hspace{0.03\textwidth}}%
+	   	\begin{subfigure}[b]{.48\textwidth}
+	\includegraphics[width=\linewidth]{figures/release} 
+	\caption{release}
    	\end{subfigure}
 
-	\begin{subfigure}[b]{.48\textwidth}
-	\includegraphics[width=\linewidth]{figures/propCache} 
-	\subcaption{ExpandPropertiesWithCache}
-   	\end{subfigure} \hspace{0.03\textwidth}  \begin{subfigure}[b]{.48\textwidth}
-	\includegraphics[width=\linewidth]{figures/release} 
-	\subcaption{release}
-   	\end{subfigure}
 \caption{Moose benchmark results piled up.}
@@
    \multirow{2}{*}{Release} 	& Sweep 		& 0606 $\pm$ 16 		&	0 $\pm$ 0 	& 0605 $\pm$ 16 	&316 $\pm$ 12 		& 1888 $\pm$ 0 	& 534 $\pm$ 58 \\
-    				    		& Compact 	& 1092 $\pm$ 02 	&	0 $\pm$ 0 	& 1092 $\pm$ 01 		&806 $\pm$ 03 			& 0193 $\pm$ 0 		& 193 $\pm$ 00 \\
+    				    		& Compact 	& 1092 $\pm$ 02 	&	0 $\pm$ 0 	& 1092 $\pm$ 01 		&806 $\pm$ 03 			& 1938 $\pm$ 0 		& 193 $\pm$ 00 \\
    \hline
@@
 
-\subsection{Sophie section}
+\subsection{Configurable benchmark}
 

### 2018-06-15 paper.tex
add a bunch of comments

@@
 Like, why growing, stable, shrinking is significant regarding garbage collection?
-same for my 3 benchs, why would I chose those 3 bench strategies?}
+same for my 3 benchs, why would I chose those 3 bench strategies?
+I know it is kind of explained in the result section but maybe it is a bit late}
 
@@
 Then, new objets with a minimum size greater than the new free chunks sizes are allocated.
-\item \emph{LargeObjects:} Only large objets with a size greater than 0xFFFF are allocated on the heap.
+\item \emph{LargeObjects:} Only large objets with a size greater than 0xFFFF \sk{put real size in Kb here} are allocated on the heap.
 \item \emph{Data0bjects:} The majority of objects that will be allocated on the heap will be Data objects, which only contains integer of a specific size (in this benchmark, one byte)
@@
 We evaluated both reference algorithms on the benchmarks built.
+\sk{shouldn't we talk about our set-ups? or maybe not because the 2 benchs were run on different computers ?}
 
@@
 
-\caption{Moose benchmark results piled up.}
+\caption{Moose benchmark results piled up (in ms).}
 \label{MooseRes1}
@@
 
-We evaluated the Moose benchmarks using the reference Mark-Compact and Mark-Sweep algorithm and by analysing the wildfly code base\footnote{http://www.wildfly.org/}.
+\sk{my benchs relies on the same data (the 2 algorithms) + same explanation of the time (full gc and so on).
+I would move the shared part in the intro of Section 4 and only keep here the part specific to the Moose bench}
+We evaluated the Moose benchmarks using the reference Mark-Compact and Mark-Sweep algorithms and by analysing the wildfly code base\footnote{http://www.wildfly.org/}.
 The results are shown in Figure \ref{MooseRes1} and \ref{MooseRes2}.
@@
 
-\todo{Sophie}
-It would be nice to evaluate both Sweeper and Planning on what we've just discussed.
+We evaluated the configurable benchmarks using the reference Mark-Compact and Mark-Sweep algorithms and by generating specific heap configurations. 
+The resulats are shown in Figure \ref{ConfigRes1} and \ref{ConfigRes2}
 
-I think we show exec time, full gc time, scavenge time, compaction time.
+\paragraph{FragmentedHeap.}
+\paragraph{LargeObjects.}
+\paragraph{Data0bjects.}
 

### 2018-06-16 paper.tex
add dacapo and start bench

@@
 
-\paragraph{ExpandPropertiesWithCache.} The ExpandPropertiesWithCache benchmark does not increase the overall memory size.
+\paragraph{ExpandPropertiesWithCache.} The ExpandProperties WithCache benchmark does not increase the overall memory size.
 Since the compacting GC has compacted the heap before starting the benchmark, there is enough room to allocate the large objects (large objects are allocated directly in old space) required by the benchmark and there is no need for a full GC during the benchmark.
@@
 \paragraph{FragmentedHeap.} 
+\begin{itemize}
+\item impact on memory size and numbers of segments
+\item impact on execution time 
+\end{itemize}
 \paragraph{LargeObjects.}
+\begin{itemize}
+\item impact on memory size 
+\item impact on number of scavenges and number of full GC's
+\end{itemize}
 \paragraph{Data0bjects.}
+\begin{itemize}
+\item impact on memory size
+\item impact on overall time (full gc, compaction time)
+\end{itemize}
 
@@
 
-\paragraph{GC benchmark suite.}
-Dacapo + see ref in Dacapo. \cite{DacapoBench}
-\todo{Sophie}
+\paragraph{GC benchmark suite.} The DaCapo benchmarks \cite{DacapoBench} is a standard benchmarking suite designed for Java applications.
+It gathers several open-source benchmarks and run them against specific metrics to assess their potential performance \sk{significance towards} on different aspects of the Java language, typically its memory management and virtual machines.
+With regard to memory management, the eleven benchmarks selected in 2006 (and updated later in 2009, to add three more benchmarks \ref{dacapowebsite} \sk{add ref in bib}) are assessed according to their number of allocated objects and their average size and their maximum number of live objects.
+The nursery (young space) survival rate is also taken into account, as well as the heap structure through object lifetime behavior.
+These criteria were chosen \sk{rephrase} as significant to asses the performance of memory management.
+
+%The part dedicated to memory management focuses on three metrics: allocation demographics, lifetime and live objects.
+The benchmarks used shows a rich and diverse object lifetime behaviors.
+--> dacapo allocates more objets, 3 time the live size of SPEC.
+%Heap volume, Heap objects, Mean object size, nursery survival rate.
+%
+%measure object size demographics (size demographics for each benchmark)
+%--> live and allocation time serie analysis
+%plot heap composition (allocations and pointer mutations)
+%--> richer lifetime behavior
 
 \paragraph{ACDC-JS.}
-ACDC-JS \cite{ACDCJS} is a configurable benchmarking tool designed to specifically measure and stress allocation and deallocation time in the JavaScript memory management model. 
+ACDC-JS \cite{ACDCJS} is a configurable \\benchmarking tool designed to specifically measure and stress allocation and deallocation time in the JavaScript memory management model. 
 It builds a heap model based on the behavior of real-time applications and aims at providing significant insights on the JavaScript memory management that cannnot be obtained from standard benchmarking suites.

### 2018-06-16 paper.tex
future works

@@
 \paragraph{Future work: more benchmarks.}
-need to add more benchmarks based on Dacapo ?
+We described two approaches to bench new memory management strategies: the Moose benchmark, based on a real-world application and the configurable benchmarking tool, highlighting edge cases by stressing the garbage collector on specific aspects.
+It would be nice to add some benchmarks based on the standard DaCapo benchmark suite so we could assess the performance of our memory management strategies via a reputed benchmarking suite.
 
 \paragraph{Future work: stress GC tests.}
-stress cases from Configurable
+We can stress the garbage collector by running the configurable benchmarking tool with specific sets of options to test the memory startegy in edge cases.
+By increasing the amount of edge-cases benchmarks led on the garbage collector, we could infer a serie of tests to secure the garbage collector behavior in these situations.
 
 \paragraph{Future work: more options.}
-\sk{maybe? I will add options in the configurator soon. Should I mention it here?}
+The number of options available in the configurable tool is still limited at the moment.
+We can take advantage of the ACDC-JS implementation, as well as some metrics used in the DaCapo paper, to get a higher number of options and improve the data collected during the benchmark.
 

### 2018-06-16 paper.tex
added setup and commetns and massage section 2.2

@@
 
+\cba{Maybe add vspace here not to have the title bottom of page}
+
 \section{Building Benchmarks}
@@
 I know it is kind of explained in the result section but maybe it is a bit late}
-
+\cba{Sure, please explain}
 To build GC benchmarks, we took two approaches.
@@
 Then, to stress the GC on specific aspects, we built a configurable benchmark that based on the configuration emulates heaps with different properties.
+
+\subsection{Set-up}
+
+To evaluate the performance, we use:
+\begin{itemize}
+	\item the production VM generated by the CI of opensmalltalk-vm\footnote{https://github.com/OpenSmalltalk/opensmalltalk-vm}, that we call Mark-Compact or Compact.
+	\item a custom compiled VM, which features a different compactor, that we call Mark-Sweep or Sweep.
+\end{itemize}
+We compiled one VMs for Linux with GCC, and one for Mac with LLVM.
+
+\paragraph{Moose benchmark.}The Moose benchmark evaluation was performed on a MacBook pro with Mac OS 10.11.6, a 2.9 Ghz processor Intel Core 5, 8 Gb 1867MHz DDR3 of RAM.
+The Mac VM was compiled using Apple LLVM version 8.0.0 (clang-800.0.38).
+The compiled VMs are 64 bits (x64) VMs.
+
+\paragraph{Configurable benchmark.}The Configurable benchmark evaluation was performed on an Asus Zenbook with Ubuntu 16.04 LTS, a 2.2 Ghz processor Intel Core 5, 8 Gb 1600MHz DDR3 of RAM.
+The Linux VM was compiled using GCC version 5.4.0.
+The compiled VMs are 32 bits (x86) VMs.
 
@@
 Running a new memory management strategy on real-world applications can give a first insight about its performance.
-Being able to configure a specific set-up, personalized environment, can allow to locally \sk{rephrase} test specific behaviors of the new implementation.
-This configurable benchmarking tool has been implemented to emulate specific heap structures and states to potentially trigger edge cases: the user has available a set of twelve options which can be sorted in two categories:
+Being able to configure a specific set-up allows to stress specific behaviors of the new implementation.
+This configurable benchmarking tool has been implemented to emulate specific heap structures and states to potentially trigger edge cases.
+The user has available a set of twelve options which can be sorted in two categories:
 
 \paragraph{Heap structure.} A first set of options allows the user to modify the heap structure, i.e.\ the amount and sizes of objects to allocate and the possible interactions with these objects.
-\todo{maybe structure is not actually a good term}
+\todo{maybe structure is not actually a good term} \cba{Heap characteristics?
+If you change change title and in previous paragraph}
 
 \begin{itemize}
-\item \emph{wantedAllocatedSize:} the cumulated size (in Kb) of objects that will be allocated
-\item \emph{minSize and maxSize:} the minimum and maximum size of objects
-\item \emph{dataObjects, weakStructures, classes, compiledMethods:} the ratio for each of these object types.
-\item \emph{accessObjects:} if set to True, access half of the objects allocated on the heap 
-\item \emph{readOnly:} if set to False, modify half of the objects allocated on the heap upon accessing  
+\item \emph{wantedAllocatedSize:} the cumulated size (in Kb) of objects that will be allocated.
+\item \emph{minSize and maxSize:} the minimum and maximum size of objects.
+\item \emph{dataObjects, weakStructures, classes, compiledMethods:} the ratio for each of these object types. \cba{Ephemerons}
+\item \emph{accessObjects:} if set to True, access half of the objects allocated on the heap .
+\item \emph{readOnly:} if set to False, modify half of the objects allocated on the heap upon accessing.
 \end{itemize}
@@
 \begin{itemize}
-\item \emph{FragmentedHeap:} First, objects are allocated one the heap, and some of them are released, leading to a potential memory fragmentation.
+\item \emph{FragmentedHeap:} First, objects are allocated on the heap, and some of them are released, leading to a potential memory fragmentation.
 Then, new objets with a minimum size greater than the new free chunks sizes are allocated.
 \item \emph{LargeObjects:} Only large objets with a size greater than 0xFFFF \sk{put real size in Kb here} are allocated on the heap.
-\item \emph{Data0bjects:} The majority of objects that will be allocated on the heap will be Data objects, which only contains integer of a specific size (in this benchmark, one byte)
+\item \emph{Data0bjects:} The majority of objects that will be allocated on the heap will be Data objects, which only contains integer of a specific size (in this benchmark, one byte) \cba{I would not say they hold integer.
+Data objects can theoretically hold float/double for example, of an int used as a bitmap}
 \end{itemize}
@@
 	\caption{ExpandPropertiesWithCache}
-	\end{subfigure\hspace{0.03\textwidth}}%
+	\end{subfigure}\hspace{0.03\textwidth}%
 	   	\begin{subfigure}[b]{.48\textwidth}

### 2018-06-16 paper.tex
new graphs

@@
 Firstly, we contacted multiple companies using the Cog VM in production on memory intensive application (>1Gb heaps) and built benchmarks out of their deployed application.
-As an example, we will discuss in the paper in Section \ref{sec:mooseBench} the Moose benchmarks implemented in collaboration with Feenk\footnote{https://feenk.com/}.
+As an example, we will discuss in the paper in Section \ref{sec:mooseBench} the Moose benchmarks implemented with the help of Feenk\footnote{https://feenk.com/}.
 Part of the business of Feenk consists in analysing software written in multiple programming languages using the open-source framework Moose~\cite{MoosePaper1,MooseBook1}.
 Using Moose, the application parses the software to analyse into a model, performs analysis on it and lastly releases the model.
+Models currently used are encoded in up to 11Gb.
 This behavior was turned into three benchmarks, growing, accessing and shrinking the heap.
@@
 Then, to stress the GC on specific aspects, we built a configurable benchmark that based on the configuration emulates heaps with different properties.
-
-\subsection{Set-up}
-
-To evaluate the performance, we use:
-\begin{itemize}
-	\item the production VM generated by the CI of opensmalltalk-vm\footnote{https://github.com/OpenSmalltalk/opensmalltalk-vm}, that we call Mark-Compact or Compact.
-	\item a custom compiled VM, which features a different compactor, that we call Mark-Sweep or Sweep.
-\end{itemize}
-We compiled one VMs for Linux with GCC, and one for Mac with LLVM.
-
-\paragraph{Moose benchmark.}The Moose benchmark evaluation was performed on a MacBook pro with Mac OS 10.11.6, a 2.9 Ghz processor Intel Core 5, 8 Gb 1867MHz DDR3 of RAM.
-The Mac VM was compiled using Apple LLVM version 8.0.0 (clang-800.0.38).
-The compiled VMs are 64 bits (x64) VMs.
-
-\paragraph{Configurable benchmark.}The Configurable benchmark evaluation was performed on an Asus Zenbook with Ubuntu 16.04 LTS, a 2.2 Ghz processor Intel Core 5, 8 Gb 1600MHz DDR3 of RAM.
-The Linux VM was compiled using GCC version 5.4.0.
-The compiled VMs are 32 bits (x86) VMs.
 
@@
 
+\paragraph{Set-up.}The Moose benchmark evaluation was performed on a MacBook pro with Mac OS 10.11.6, a 2.9 Ghz processor Intel Core 5, 8 Gb 1867MHz DDR3 of RAM.
+The Mac VM was compiled using Apple LLVM version 8.0.0 (clang-800.0.38).
+The compiled VMs are 64 bits (x64) VMs.
+
+%To evaluate the performance, we use:
+%\begin{itemize}
+%	\item the production VM generated by the CI of opensmalltalk-vm\footnote{https://github.com/OpenSmalltalk/opensmalltalk-vm}, that we call Mark-Compact or Compact.
+%	\item a custom compiled VM, which features a different compactor, that we call Mark-Sweep or Sweep.
+%\end{itemize}
+
 %
@@
 
+\paragraph{Set-up.}The Configurable benchmark evaluation was performed on an Asus Zenbook with Ubuntu 16.04 LTS, a 2.2 Ghz processor Intel Core 5, 8 Gb 1600MHz DDR3 of RAM.
+The Linux VM was compiled using GCC version 5.4.0.
+The compiled VMs are 32 bits (x86) VMs.
+
 We evaluated the configurable benchmarks using the reference Mark-Compact and Mark-Sweep algorithms and by generating specific heap configurations. 
-The resulats are shown in Figure \ref{ConfigRes1} and \ref{ConfigRes2}
+The results are shown in Figure \ref{ConfigRes1} and \ref{ConfigRes2}
 

### 2018-06-16 paper.tex
added table

@@
 	\includegraphics[width=\linewidth]{figures/load} 
-	\caption{Load\vspace{0.5cm}}
+	\caption{Load\vspace{0.2cm}}
    	\end{subfigure}\hspace{0.03\textwidth}% 
@@
 	\includegraphics[width=\linewidth]{figures/prop} 
-	\caption{ExpandProperties\vspace{0.5cm}}
+	\caption{ExpandProperties\vspace{0.2cm}}
    	\end{subfigure}	
@@
 \begin{figure*}[thb]
-\begin{tabular}{|c|c|c|c|c|c|c|c|}
+\begin{tabular}{|c|c|l|l|l|l|l|l|}
    \hline
@@
    \hline
-   \multirow{2}{*}{Load} & Sweep 	& 159821 $\pm$ 2218 &	73599 $\pm$ 1281 	& 10863 $\pm$ 726 	&1398 $\pm$ 091 	& 193 $\pm$ 0 	& 959 $\pm$ 9.69 \\
-    				    & Compact 	& 162305 $\pm$ 2790 &	72150 $\pm$ 0782 	& 14091 $\pm$ 972 	&4661 $\pm$ 289 	& 193 $\pm$ 0 	& 909 $\pm$ 9.69 \\
+   \multirow{4}{*}{Load} & Sweep 	& 159821 $\pm$ 2218 &	73599 $\pm$ 1281 	& 10863 $\pm$ 726 	&1398 $\pm$ 91 	& 193 $\pm$ 0 	& 959 $\pm$ 9.69 \\
+    				    & Compact 	& 162305 $\pm$ 2790 &	72150 $\pm$ 782 	& 14091 $\pm$ 972 	&4661 $\pm$ 289 	& 193 $\pm$ 0 	& 909 $\pm$ 9.69 \\
+    				    & Compact-16 	& 96043 $\pm$ 540 &	14279 $\pm$ 124 	& 12440 $\pm$ 217 	&5575 $\pm$ 36 	& 283 $\pm$ 0 	& 1055 $\pm$ 0 \\
+    				    & Sweep-16 		& 91721 $\pm$ 1131 &	13862 $\pm$ 49 	& 9352 $\pm$ 73 	&1208 $\pm$ 9 	& 283 $\pm$ 0 	& 988 $\pm$ 0 \\
    \hline
-   \multirow{2}{*}{ExpandProperties} 	& Sweep 		& 232940 $\pm$ 01287 &	55497 $\pm$ 0238 	& 12240 $\pm$ 285 	&1178 $\pm$ 6.66 	& 959 $\pm$ 9.69 & 1888 $\pm$ 0 \\
-    				    			& Compact 	& 183314 $\pm$ 13090 &	53165 $\pm$ 2343 	& 12496 $\pm$ 197 	&3775 $\pm$ 0108 	& 909 $\pm$ 9.69 & 1938 $\pm$ 0 \\
+   \multirow{4}{*}{ExpandProperties} 	& Sweep 		& 232940 $\pm$ 1287 &	55497 $\pm$ 238 	& 12240 $\pm$ 285 	&1178 $\pm$ 6.66 	& 959 $\pm$ 9.69 & 1888 $\pm$ 0 \\
+    				    			& Compact 	& 183314 $\pm$ 13090 &	53165 $\pm$ 2343 	& 12496 $\pm$ 197 	&3775 $\pm$ 108 	& 909 $\pm$ 9.69 & 1938 $\pm$ 0 \\
+    				    & Compact-16 	& 100315 $\pm$ 138 &	9372 $\pm$ 75 	& 9466 $\pm$ 87 	&3301 $\pm$ 31 	& 1055 $\pm$ 0 	& 2028 $\pm$ 0 \\
+    				    & Sweep-16 		& 107914 $\pm$ 216 &	9714 $\pm$ 34 	& 9384 $\pm$ 107 	&864 $\pm$ 5 	& 988 $\pm$ 0 	& 2011 $\pm$ 0 \\
    \hline
-   ExpandProperties 		& Sweep 		& 510013 $\pm$ 05812 &	64234 $\pm$ 0823 	& 23225 $\pm$ 100 	&2011 $\pm$ 10 	& 1888 $\pm$ 0 	& 1888 $\pm$ 0 \\
-   WithCache  			& Compact 	& 408169 $\pm$ 20444 &	64492 $\pm$ 1234 	& 0 $\pm$ 0 		&0 $\pm$ 0 		& 1938 $\pm$ 0 		& 1938 $\pm$ 0 \\
+    					& Sweep 		& 510013 $\pm$ 5812 	&	64234 $\pm$ 823 	& 23225 $\pm$ 100 	&2011 $\pm$ 10 	& 1888 $\pm$ 0 	& 1888 $\pm$ 0 \\
+     	ExpandProperties	& Compact 	& 408169 $\pm$ 20444 	&	64492 $\pm$ 1234 	& 0 $\pm$ 0 		&0 $\pm$ 0 		& 1938 $\pm$ 0 	& 1938 $\pm$ 0 \\
+    		WithCache	& Compact-16 	& 105389 $\pm$ 1067 	&	6467 $\pm$ 47 		& 3019 $\pm$ 24 	&937 $\pm$ 6 		& 2028 $\pm$ 0 	& 2028 $\pm$ 0 \\
+    				    	& Sweep-16 	& 114325 $\pm$ 403 	&	6323 $\pm$ 23 		& 2347 $\pm$ 13 	&209 $\pm$ 6 		& 2011 $\pm$ 0 	& 2011 $\pm$ 0 \\
    \hline
-   \multirow{2}{*}{Release} 	& Sweep 		& 0606 $\pm$ 16 		&	0 $\pm$ 0 	& 0605 $\pm$ 16 	&316 $\pm$ 12 		& 1888 $\pm$ 0 	& 534 $\pm$ 58 \\
-    				    		& Compact 	& 1092 $\pm$ 02 	&	0 $\pm$ 0 	& 1092 $\pm$ 01 		&806 $\pm$ 03 			& 1938 $\pm$ 0 		& 193 $\pm$ 00 \\
+   \multirow{4}{*}{Release} 	& Sweep 		& 606 $\pm$ 16 		&	0 $\pm$ 0 	& 605 $\pm$ 16 	&316 $\pm$ 12 		& 1888 $\pm$ 0 		& 534 $\pm$ 58 \\
+    				    		& Compact 	& 1092 $\pm$ 2 		&	0 $\pm$ 0 	& 1092 $\pm$ 1 	&806 $\pm$ 3 			& 1938 $\pm$ 0 		& 193 $\pm$ 0 \\
+    				    & Compact-16 	& 1120 $\pm$ 13 		&	0 $\pm$ 0 		& 1119 $\pm$ 13 	&829 $\pm$ 8 			& 2028 $\pm$ 0 		& 300 $\pm$ 0 \\
+    				    & Sweep-16 		& 610 $\pm$ 1 			&	0 $\pm$ 0 		& 610 $\pm$ 1 		&314 $\pm$ 2 			& 2011 $\pm$ 0 		& 535 $\pm$ 203 \\
    \hline

### 2018-06-16 paper.tex
finished 16

@@
 
+We evaluated the Moose benchmarks using the reference Mark-Compact and Mark-Sweep algorithms and by analysing the wildfly code base\footnote{http://www.wildfly.org/}.
+We distinguish four configurations:
+\begin{itemize}
+\item \emph{Sweep}: VM with standard GC parameters and the Mark-Sweep algorithm.
+\item \emph{Compact}: VM with standard GC parameters and the Mark-Compact algorithm.
+\item \emph{Compact-16}: VM with standard GC parameters, except Eden size which is 16 times larger, and the Mark-Compact algorithm.
+\item \emph{Sweep-16}: VM with standard GC parameters, except Eden size which is 16 times larger, and the Mark-Sweep algorithm.
+\end{itemize}
+With a larger eden, more overall memory is used (in our case 90 extra Mb are used), but scavenge time is expected to be lower.
+Indeed, the scavenger is an implementation of Ungar's scavenger~\cite{Scavenger} and the scavenge time is proportional to the number of surviving objects.
 %To evaluate the performance, we use:
@@
 I would move the shared part in the intro of Section 4 and only keep here the part specific to the Moose bench}
-We evaluated the Moose benchmarks using the reference Mark-Compact and Mark-Sweep algorithms and by analysing the wildfly code base\footnote{http://www.wildfly.org/}.
 The results are shown in Figure \ref{MooseRes1} and \ref{MooseRes2}.
@@
 Total execution time is the total execution time, including scavenge and full GC time.
+Scavenge time is drastically smaller with a larger Eden, leading to overall better performance.
 
-\paragraph{Load.} The Load benchmark grows the memory by 750 Mb with Sweep and by 700 Mb with Compact.
+\paragraph{Load.} The Load benchmark grows the memory by around 700Mb.
 Compaction time is smaller for the sweep algorithm, making the benchmark a little bit faster in the Mark-Sweep, which is to be expected since Sweep does not need to waste time updating pointers because objects are not moved in memory.
@@
 We believe this is due to worse locality of objects.
+With a larger Eden, the scavenge time is much smaller and the execution time is a bit faster due to improved object locality.
 
@@
 In addition, execution time is worse, as for the previous benchmark.
+With a larger Eden, the scavenge time is much smaller and the execution time is a way faster due to improved object locality.
 
@@
 Compaction time is much faster with Sweep, but as we have discussed, for less efficiency.
+Changing Eden size has not really any impact here, since it's essentially a pure full GC benchmark.
 

### 2018-06-20 paper.tex
Eliot's pass

@@
 \newcommand{\cba}[1]{\color{purple}\fbox{\bfseries\sffamily\scriptsize Clement:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
+\newcommand{\eem}[1]{\color{green}\fbox{\bfseries\sffamily\scriptsize Eliot:}{\sf\small$\blacktriangleright$\textit{#1}$\blacktriangleleft$}\color{black}}
+
 %\newcommand*{rotatebox{75}}
@@
   %\position{Position1}
-  \department{Master student}              %% \department is recommended
+  \department{Masters student}              %% \department is recommended
   \institution{Universit\'e de Lille 1}            %% \institution is required
@@
 \begin{abstract}
-One of the next step to improve the Cog virtual machine, the default virtual machine for multiple programming languages such as Pharo, Squeak and Newspeak, is to decrease the garbage collection pause time.
-A benchmarking infrastructure and reference garbage collection algorithm implementations are required to evaluate the performance of a new algorithm and compare it.
-Cog features a Mark-Compact algorithm, used in production, and we introduced a Mark-Sweep algorithm as the two reference algorithms.
+One of the next steps to improve the Cog virtual machine, the default virtual machine for multiple programming languages in the Smalltalk family, such as Pharo, Squeak and Newspeak, is to decrease garbage collection pause times.
+Reference garbage collection algorithm implementations and a benchmarking infrastructure are required to evaluate the performance of a new algorithm and compare it.
+Cog features a Mark-Compact algorithm, used in production, to which we added a Mark-Sweep algorithm, providing two reference algorithms.
 Benchmarks are built using two different approaches.
@@
 Secondly, we built a configurable benchmark which simulates an application with different heap properties to be able to stress specific aspects of the memory management.
-We then evaluated the two reference algorithms on the infrastructure built to have reference benchmark results.
+We then evaluated the two reference algorithms on the infrastructure built to obtain reference benchmark results.
 \end{abstract}
@@
 
-Building and tuning a new GC algorithm requires to evaluate its behavior. A benchmarking infrastructure is required to do so. To build GC benchmarks, we took two approaches.
+Building and tuning a new GC algorithm requires evaluating its behavior. A benchmarking infrastructure is required to do so. To build GC benchmarks, we took two approaches.
 
-Firstly, we contacted multiple companies using the Cog VM in production on memory intensive application (>1Gb heaps) and built benchmarks out of their deployed application.
+First, we contacted multiple companies looking for use of the Cog VM in production with memory intensive applications (>1Gb heaps) and built benchmarks out of one of their deployed applications.
 As an example, we will discuss in the paper in Section \ref{sec:mooseBench} the Moose benchmarks implemented with the help of Feenk\footnote{https://feenk.com/}.
@@
 
-Secondly, we implemented a configurable benchmark: the idea is to set-up specific heap structures that will stress the garbage collector.
-A set of specific options is available to the user to tune the type of allocated objects and some of their features.
+Second, we implemented a configurable benchmark: the idea is to set-up specific heap structures that will stress the garbage collector.
+A set of options is available to the user to tune the type of allocated objects and some of their features.
 These options can be associated with different heap states: growing, stable or shrinking, like in our other approach.
@@
 To evaluate a new algorithm, we also need to compare it to reference implementations.
-The Cog VM features a Mark-Compact algorithm in production.
-The compaction phase is quite specific since it moves objects in memory and spend time updating references.
-In addition to this algorithm, we implemented a Mark-Sweep, which does not move objects in memory and is quicker to perform.
+The Cog VM features a Mark-Compact algorithm used in production.
+The compaction phase is quite specific since it slides objects down in memory while handling pinned objects and spends time updating references to point to their eventual locations.
+In addition to this algorithm, we implemented a Mark-Sweep, which does not move objects in memory and is thus quicker to perform.
 Section \ref{sec:ref} details the two implementations.
 
-We evaluated both algorithm on the benchmarks we built to provide results of the reference algorithms on the infrastructure built.
+We evaluated both algorithms on the benchmarks we built to provide results of the reference algorithms on the infrastructure built.
 
@@
 
-Part of the business of Feenk consists in analyzing softwares.
+Part of the business of Feenk consists in analyzing large software systems.
 To do so, the application parses a mse file into a model, performs some analysis, and then releases the model and generates analysis results.
@@
 
-We built four different benchmarks that falls into three categories: growing, stable and shrinking heaps.
+We built four different benchmarks that fall into three categories: growing, stable and shrinking heaps.
 The benchmarks were built on top of the stable Moose image\footnote{http://www.moosetechnology.org/}.
@@
 \begin{itemize}
-	\item \emph{LoadFromMSE}: parses a mse file into a moose model, effectively loading a graph of object. This is the only benchmark not taking a moose model as a parameter.
-	\item \emph{ExpandProperties}: computes all interesting properties of the model, discard properties rarely used or consuming too much memory, cache common properties not using too much memory.
-It effectively loads a graph of object connected in many points to the original graph.
+	\item \emph{LoadFromMSE}: parses a mse file into a moose model, effectively loading a graph of objects. This is the only benchmark not taking a moose model as a parameter.
+	\item \emph{ExpandProperties}: computes all interesting properties of the model, discarding properties rarely used or consuming too much memory, cacheing common properties not using too much memory.
+It effectively loads a graph of objects connected in many points to the original graph.
 \end{itemize}
 
-\paragraph{Stable heap.} The stable heap benchmarks perform random memory accesses in a heap which size remains approximately the same.
+\paragraph{Stable heap.} The stable heap benchmarks perform random memory accesses in a heap whose size remains approximately the same.
 We built one benchmark, \emph{ExpandPropertiesWithCache}, which computes all interesting properties of the model, common properties with low memory footprint are already cached and other properties are discarded once computed.
@@
 
-Running a new memory management strategy on real-world applications can give a first insight about its performance.
-Being able to configure a specific set-up allows to stress specific behaviors of the new implementation.
-This configurable benchmarking tool has been implemented to emulate specific heap structures and states to potentially trigger edge cases.
-The user has available a set of twelve options which can be sorted in two categories:
+Running a new memory management strategy on real-world applications can give an early insight into its performance.
+Being able to configure a specific set-up allows stressing specific behaviors of the new implementation.
+This configurable benchmarking tool has been implemented to emulate specific heap structures and states potentially to trigger edge cases.
+The user has available a set of twelve options which can be sorted into two categories:
 
@@
 \begin{itemize}
-\item \emph{wantedAllocatedSize:} the cumulated size (in Kb) of objects that will be allocated.
-\item \emph{minSize and maxSize:} the minimum and maximum size of objects.
+\item \emph{wantedAllocatedSize:} the cumulative size (in Kb) of objects that will be allocated.
+\item \emph{minSize and maxSize:} the minimum and maximum sizes of objects.
 \item \emph{dataObjects, weakStructures, classes, compiledMethods:} the ratio for each of these object types. \cba{Ephemerons}
@@
 \paragraph{Heap state.} The heap state can also be tuned.
-Once objects have been allocated on the heap according to the selected options, it is possible whether to allocate more objects (growing heap), to release objects that have been previously allocated (shrinking heap), or to stay in the current state and potentially accessing and modifying objects that have been previously allocated (stable heap).
-These options echoes the 3 benchmarks depicted in Section \ref{sec:mooseBench}\\
+Once objects have been allocated on the heap according to the selected options, it is possible to decide whether to allocate more objects (growing heap), to release objects that have been previously allocated (shrinking heap), or to stay in the current state and potentially access and modify objects that have been previously allocated (stable heap).
+These options echo the 3 benchmarks depicted in Section \ref{sec:mooseBench}\\
 
-We build three benchmarks out of the potential configurations:
+We built three benchmarks out of the potential configurations:
 \begin{itemize}
-\item \emph{FragmentedHeap:} First, objects are allocated on the heap, and some of them are released, leading to a potential memory fragmentation.
-Then, new objets with a minimum size greater than the new free chunks sizes are allocated.
-\item \emph{LargeObjects:} Only large objets with a size greater than 0xFFFF \sk{put real size in Kb here} are allocated on the heap.
-\item \emph{Data0bjects:} The majority of objects that will be allocated on the heap will be Data objects, which only contains integer of a specific size (in this benchmark, one byte) \cba{I would not say they hold integer.
+\item \emph{FragmentedHeap:} First, objects are allocated on the heap, and some of them are released, leading to potential memory fragmentation.
+Then, new objects with a minimum size greater than the new free chunks sizes are allocated.
+\item \emph{LargeObjects:} Only large objects with a size greater than 0xFFFF \sk{put real size in Kb here} are allocated on the heap.
+\item \emph{Data0bjects:} The majority of objects that will be allocated on the heap will be Data objects, which only contains integers of a specific size (in this benchmark, one byte) \cba{I would not say they hold integer.
 Data objects can theoretically hold float/double for example, of an int used as a bitmap}
@@
 It uses the fact that there is room for a forwarding pointer in all objects to store the eventual position of an object in the first field.
-It therefore first locates a large free chunk, or eden or a memory segment, to use as the savedFirstFieldsSpace, which it uses to store the first fields of objects that will be compacted.
+It therefore first locates a large free chunk, or eden or a memory segment, to use as the savedFirstFieldsSpace, which it uses to store the first fields of objects that will be compacted, these fields being overwritten by each object's eventual position.
 It then makes at least three passes through the heap.
 
-The first pass plans where live movable objects will go, copying their first field to the next slot in savedFirstFieldsSpace, and setting their forwarding pointer to point to their eventual location (see planCompactSavingForwarders).
-The second pass updates all pointers in live pointer objects to point to objects' final destinations, including the fields in savedFirstFieldsSpace (see updatePointers and updatePointersInMobileObjects).
-The third pass moves objects to their final positions, unmarking objects, and restoring saved first fields as it does so (see copyAndUnmark: and copyAndUnmarkMobileObjects).
+The first pass plans where live movable objects will go, copying the first field to the next slot in savedFirstFieldsSpace, and setting the forwarding pointer to point to the object's eventual location.
+The second pass updates all pointers in live pointer objects to point to objects' final destinations, including the fields in savedFirstFieldsSpace.
+The third pass moves objects to their final positions, unmarking objects, and restoring saved first fields as it does so.
 If the forwarding fields of live objects in the to-be-moved portion of the entire heap won't fit in savedFirstFieldsSpace, then additional passes can be made until the entire heap has been compacted.
-When snapshotting multiple passes are made, but when doing a normal GC only one pass is made.
+When snapshotting (saving the system to a file for later resumption, see below) multiple passes are made, but when doing a normal GC only a single pass is made.
 
-Each pass uses a three finger algorithm, a simple extension of the classic two finger algorithm with an extra finger used to identify the lowest pinned object between the to and from fingers.
+Each pass uses a three finger algorithm, a simple extension of the classic two finger algorithm with an extra middle finger used to refer to the lowest pinned object, if any, between the to and from fingers.
 Objects are moved down, starting at the first free object or chunk, provided that they fit below the lowest pinned object above the to finger.
 When an object won't fit the to finger is moved above the pinned object and the third finger is reset to the next pinned object below the from finger, if any.
+The scheme preserves object order amongst movable objects, making no attempt to fit objects in any gaps left below pinned objects.
 
@@
 
-Cog's Mark-Compact algorithm, named \emph{SpurSweeper}, is a sweep-only algorithm, setting the compactor to SpurSweeper effectively changes the fullGC to a mark-sweep non-moving algorithm.
-It iterates over all old space in linear order.
+Cog's Mark-Sweep algorithm, named \emph{SpurSweeper}, is a sweep-only algorithm.
+Setting the compactor to SpurSweeper effectively changes the fullGC to a mark-sweep non-moving algorithm.
+It iterates over all objects in old space in linear order.
 Each time an unmarked object or a free chunk is met, it coalesces it with subsequent ones and updates the free lists with a new larger free chunk.
@@
 The heap, when collected by the Mark-Sweep, has a lot of free chunks when the heap size varies a lot leading to a large snapshot.
-However, non compacting have also interesting advantages (lower garbage collection pause time for example).
+However, the non compacting algorithm also has interesting advantages, for  example lower garbage collection pause times.
 
-To be able to have a garbage collector with a low pause time and still be able to perform efficiently snapshots, we design a hybrid compactor solution.
-The VM tells the compactor if the garbage collection is performed for snapshot or not, and the compactor chooses to apply one algorithm or the other to have low compaction pause time or to snapshot with a low memory footprint.
+To be able to have a garbage collector with a low pause time and still be able to perform snapshots efficiently, we designed a hybrid compactor solution which includes both SpurSelectiveCompactor and SpurPlanningCompactor.
+The VM tells the hybrid compactor if the garbage collection is to be performed for snapshot or not, and the compactor chooses to apply one algorithm or the other to achieve either a low compaction pause time or a snapshot with a minimised memory footprint.
 
@@
 
-\paragraph{Set-up.}The Moose benchmark evaluation was performed on a MacBook pro with Mac OS 10.11.6, a 2.9 Ghz processor Intel Core 5, 8 Gb 1867MHz DDR3 of RAM.
+\paragraph{Set-up.}The Moose benchmark evaluation was performed on a MacBook pro running Mac OS 10.11.6, with a 2.9 Ghz Intel Core i5 processor, and 8 Gb of  1867MHz DDR3 RAM.
 The Mac VM was compiled using Apple LLVM version 8.0.0 (clang-800.0.38).
-The compiled VMs are 64 bits (x64) VMs.
+The compiled VMs are native 64 bit applications with 64-bit objects and pointers.
 
@@
 In the results, the compaction time includes the time spent in the compact phase in the Mark-Compact and in the sweep phase in the Mark-Sweep.
-Total full GC time is the total time spent in full GCs, including the compaction time.
+Total full GC time is the total time spent in full GCs, including  compaction time.
 Scavenge time is the time spent in scavenges, \emph{i.e.,} garbage collection of young objects only.
 Total execution time is the total execution time, including scavenge and full GC time.
-Scavenge time is drastically smaller with a larger Eden, leading to overall better performance.
+Scavenge time is drastically smaller with a larger Eden, leading to better overall performance.
 
@@
 
-\paragraph{ExpandPropertiesWithCache.} The ExpandProperties WithCache benchmark does not increase the overall memory size.
+\paragraph{ExpandPropertiesWithCache.} The ExpandPropertiesWithCache benchmark does not increase the overall memory size.
 Since the compacting GC has compacted the heap before starting the benchmark, there is enough room to allocate the large objects (large objects are allocated directly in old space) required by the benchmark and there is no need for a full GC during the benchmark.
-However, with the sweep algorithm, there are multiple small available memory chunks which are not big enough to hold all the requested large objects, so a full GC is required.
+However, with the sweep algorithm, there are multiple small free memory chunks which are not big enough to hold all the requested large objects, so a full GC is required.
 In addition, execution time is worse, as for the previous benchmark.
-With a larger Eden, the scavenge time is much smaller and the execution time is a way faster due to improved object locality.
+With a larger Eden, the scavenge time is much smaller and the execution time is much faster due to improved object locality.
 
@@
 With the compacting GC, memory goes back to its original size.
-With Sweep, multiple memory segments are left with a few live objects, forbidding the OS to reclaim those segments.
+With Sweep, multiple memory segments are left with a few live objects, preventing the Vm from returning those segments to the OS.
 Hence, the memory decreases drastically, but not to the original size.
 Compaction time is much faster with Sweep, but as we have discussed, for less efficiency.
-Changing Eden size has not really any impact here, since it's essentially a pure full GC benchmark.
+Changing Eden size has not much, if any, impact here, since it's essentially a pure full GC benchmark.
 
@@
 
-\paragraph{Set-up.}The Configurable benchmark evaluation was performed on an Asus Zenbook with Ubuntu 16.04 LTS, a 2.2 Ghz processor Intel Core 5, 8 Gb 1600MHz DDR3 of RAM.
+\paragraph{Set-up.}The Configurable benchmark evaluation was performed on an Asus Zenbook running Ubuntu 16.04 LTS, with a 2.2 Ghz processor Intel Core i5, and 8 Gb of 1600MHz DDR3 RAM.
 The Linux VM was compiled using GCC version 5.4.0.
-The compiled VMs are 32 bits (x86) VMs.
+The compiled VMs are 32 bit (x86) applications with 32-bit objects and pointers \footnote{Spur uses a common object header for both 32-bit and 64-bit versions, so 32-bit objects still have a 64-bit header and are all aligned on a 64-bit boundary.}
 
@@
 
-\paragraph{GC benchmark suite.} The DaCapo benchmarks \cite{DacapoBench} is a standard benchmarking suite designed for Java applications.
-It gathers several open-source benchmarks and run them against specific metrics to assess their potential performance \sk{significance towards} on different aspects of the Java language, typically its memory management and virtual machines.
-With regard to memory management, the eleven benchmarks selected in 2006 (and updated later in 2009, to add three more benchmarks \ref{dacapowebsite} \sk{add ref in bib}) are assessed according to their number of allocated objects and their average size and their maximum number of live objects.
-The nursery (young space) survival rate is also taken into account, as well as the heap structure through object lifetime behavior.
-These criteria were chosen \sk{rephrase} as significant to asses the performance of memory management.
+\paragraph{GC benchmark suite.} The DaCapo benchmark suite \cite{DacapoBench} is a standard set of benchmarks designed for Java applications.
+It gathers together several open-source benchmarks and runs them against specific metrics to assess their potential performance \sk{significance towards} on different aspects of the Java language implementation, typically its memory management and virtual machines.
+With regard to memory management, the eleven benchmarks selected in 2006 (and updated later in 2009 to add three more benchmarks \ref{dacapowebsite} \sk{add ref in bib}) are assessed according to the number of allocated objects, their average size, and the maximum number of live objects.
+The nursery (new space) survival rate is also taken into account, as well as the heap structure through object lifetime behavior.
+These criteria were identified as significant in assessing the performance of memory management \eem{by whom?}.
 
@@
 \paragraph{ACDC-JS.}
-ACDC-JS \cite{ACDCJS} is a configurable \\benchmarking tool designed to specifically measure and stress allocation and deallocation time in the JavaScript memory management model. 
-It builds a heap model based on the behavior of real-time applications and aims at providing significant insights on the JavaScript memory management that cannnot be obtained from standard benchmarking suites.
+ACDC-JS \cite{ACDCJS} is a configurable \\benchmarking tool designed to measure and stress allocation and deallocation time in the JavaScript memory management model. 
+It builds a heap model based on the behavior of real-time \eem{real world?} applications and aims at providing significant insights into JavaScript memory management that cannot be obtained from standard benchmarking suites.
 The value of specific heap options can be set by the user.
@@
 
+\paragraph{Future work: incremental GC.}
+This investigation, including as it does a framework for compaction that includes a new store check, helps us in designing a fully incremental Mark-Sweep-Compact algorithm with very low pause times, suitable for interactive applications such as animation and gaming.
+
 %% Bibliography

### 2018-06-20 paper.tex
took most of the comments into account

@@
 
-Second, we implemented a configurable benchmark: the idea is to set-up specific heap structures that will stress the garbage collector.
+Second, we implemented a configurable benchmark: the idea is to set-up specific heap characteristics that will stress the garbage collector.
 A set of options is available to the user to tune the type of allocated objects and some of their features.
 These options can be associated with different heap states: growing, stable or shrinking, like in our other approach.
-\todo{fin a ref}
-\sk{add the case we benchmarked with this approach}
+We ran three benchmark cases with this tool: fragmented heap, large objects in heap and majority of data objects in heap.
 This implementation and the use-cases are detailed in Section \ref{sec:confBench}
@@
 Being able to configure a specific set-up allows stressing specific behaviors of the new implementation.
-This configurable benchmarking tool has been implemented to emulate specific heap structures and states potentially to trigger edge cases.
+This configurable benchmarking tool has been implemented to emulate specific heap characteristics and states potentially to trigger edge cases.
 The user has available a set of twelve options which can be sorted into two categories:
 
-\paragraph{Heap structure.} A first set of options allows the user to modify the heap structure, i.e.\ the amount and sizes of objects to allocate and the possible interactions with these objects.
-\todo{maybe structure is not actually a good term} \cba{Heap characteristics?
-If you change change title and in previous paragraph}
+\paragraph{Heap characteristics.} A first set of options allows the user to modify the heap characteristics, i.e.\ the amount and sizes of objects to allocate and the possible interactions with these objects.
 
@@
 \item \emph{minSize and maxSize:} the minimum and maximum sizes of objects.
-\item \emph{dataObjects, weakStructures, classes, compiledMethods:} the ratio for each of these object types. \cba{Ephemerons}
+\item \emph{dataObjects, weakStructures, classes, compiledMethods, ephemerons:} the ratio for each of these object types.
 \item \emph{accessObjects:} if set to True, access half of the objects allocated on the heap .
@@
 \item \emph{LargeObjects:} Only large objects with a size greater than 0xFFFF \sk{put real size in Kb here} are allocated on the heap.
-\item \emph{Data0bjects:} The majority of objects that will be allocated on the heap will be Data objects, which only contains integers of a specific size (in this benchmark, one byte) \cba{I would not say they hold integer.
-Data objects can theoretically hold float/double for example, of an int used as a bitmap}
+\item \emph{Data0bjects:} The majority of objects that will be allocated on the heap will be Data objects, which will only contains one-byte-sized integers of a specific size in this benchmark  
 \end{itemize}
@@
 We evaluated both reference algorithms on the benchmarks built.
-\sk{shouldn't we talk about our set-ups? or maybe not because the 2 benchs were run on different computers ?}
+
+In the results, the compaction time includes the time spent in the compact phase in the Mark-Compact and in the sweep phase in the Mark-Sweep.
+Total full GC time is the total time spent in full GCs, including  compaction time.
+Scavenge time is the time spent in scavenges, \emph{i.e.,} garbage collection of young objects only.
+Total execution time is the total execution time, including scavenge and full GC time.
+Scavenge time is drastically smaller with a larger Eden, leading to better overall performance.
 
@@
 	\includegraphics[width=\linewidth]{figures/release} 
-	\caption{release}
+	\caption{Release}
    	\end{subfigure}
@@
 
-\sk{my benchs relies on the same data (the 2 algorithms) + same explanation of the time (full gc and so on).
-I would move the shared part in the intro of Section 4 and only keep here the part specific to the Moose bench}
 The results are shown in Figure \ref{MooseRes1} and \ref{MooseRes2}.
-In the results, the compaction time includes the time spent in the compact phase in the Mark-Compact and in the sweep phase in the Mark-Sweep.
-Total full GC time is the total time spent in full GCs, including  compaction time.
-Scavenge time is the time spent in scavenges, \emph{i.e.,} garbage collection of young objects only.
-Total execution time is the total execution time, including scavenge and full GC time.
-Scavenge time is drastically smaller with a larger Eden, leading to better overall performance.
 
@@
 
-\paragraph{Set-up.}The Configurable benchmark evaluation was performed on an Asus Zenbook running Ubuntu 16.04 LTS, with a 2.2 Ghz processor Intel Core i5, and 8 Gb of 1600MHz DDR3 RAM.
+\paragraph{Set-up.}The Configurable benchmark evaluation was performed on an Asus Zenbook running Ubuntu 18.04 LTS, with a 2.2 Ghz processor Intel Core i5, and 8 Gb of 1600MHz DDR3 RAM.
 The Linux VM was compiled using GCC version 5.4.0.
@@
 \paragraph{FragmentedHeap.} 
+allocate n objects of size x, release half of them, and reallocate n/2 objects of size y>x.
+Compare the performance (time spent in full GC, scavenges, compacting, and execution time) for both Mark-Sweep and Mark-Compact
 \begin{itemize}
@@
 \paragraph{LargeObjects.}
+allocate objects larger than 0xffff. Compare two VMs with different number of segments.
 \begin{itemize}
@@
 \paragraph{Data0bjects.}
+ allocate a lot of data objects (90\% of all allocated objects during the bench). Compare for Mark-Sweep and Mark-Compact
 \begin{itemize}
@@
 \paragraph{GC benchmark suite.} The DaCapo benchmark suite \cite{DacapoBench} is a standard set of benchmarks designed for Java applications.
-It gathers together several open-source benchmarks and runs them against specific metrics to assess their potential performance \sk{significance towards} on different aspects of the Java language implementation, typically its memory management and virtual machines.
-With regard to memory management, the eleven benchmarks selected in 2006 (and updated later in 2009 to add three more benchmarks \ref{dacapowebsite} \sk{add ref in bib}) are assessed according to the number of allocated objects, their average size, and the maximum number of live objects.
+It gathers together several open-source benchmarks and runs them against specific metrics to assess their potential significance towards different aspects of the Java language implementation, typically its memory management and virtual machines.
+With regard to memory management, the eleven benchmarks selected in 2006 (and updated later in 2009 to add three more benchmarks \cite{DaCapo}) are assessed according to the number of allocated objects, their average size, and the maximum number of live objects.
 The nursery (new space) survival rate is also taken into account, as well as the heap structure through object lifetime behavior.
@@
 ACDC-JS \cite{ACDCJS} is a configurable \\benchmarking tool designed to measure and stress allocation and deallocation time in the JavaScript memory management model. 
-It builds a heap model based on the behavior of real-time \eem{real world?} applications and aims at providing significant insights into JavaScript memory management that cannot be obtained from standard benchmarking suites.
+It builds a heap model based on the behavior of real-world applications and aims at providing significant insights into JavaScript memory management that cannot be obtained from standard benchmarking suites.
 The value of specific heap options can be set by the user.

### 2018-06-20 paper.tex
Minor abstract simplification.

@@
 \begin{abstract}
-One of the next steps to improve the Cog virtual machine, the default virtual machine for multiple programming languages in the Smalltalk family, such as Pharo, Squeak and Newspeak, is to decrease garbage collection pause times.
+One of the next steps to improve Cog, the default virtual machine for multiple programming languages in the Smalltalk family, such as Pharo, Squeak and Newspeak, is to decrease garbage collection pause times.
 Reference garbage collection algorithm implementations and a benchmarking infrastructure are required to evaluate the performance of a new algorithm and compare it.

### 2018-06-20 paper.tex
Move labels to their section lines, othewrwise they don't label anything and the \ref{foo} entries don't work.
Add a reference for the Spur paper in ISMM.  Reference it at the end of the first para in Sec 3, "Reference Implementations".

@@
 
-\keywords{Benchmark, Garbage Collector, Virtual machine, Managed runtime}  %% \keywords are mandatory in final camera-ready submission
+\keywords{Benchmark, Garbage Collector, Virtual Machine, Managed Runtime}  %% \keywords are mandatory in final camera-ready submission
 
@@
 
-\section{Introduction}
-\label{sec:intro}
+\section{Introduction} \label{sec:intro}
 
 The Cog virtual machine (VM), the default VM for multiple programming languages such Pharo \cite{PharoByExample}, Squeak \cite{SqueakByExample} and Newspeak \cite{NewspeakOopsla}, currently features in production a stop-the-world Mark-Compact algorithm as the full garbage collector (GC) algorithm.
-The algorithm has a high throughput but a high pause time during which the application is not responsive (the GC interrupts the application for multiple seconds on modern Macbooks starting from multiple Gbs heaps).
+The algorithm has a high throughput but a high pause time during which the application is not responsive (the GC interrupts the application for multiple seconds on modern Macbooks when using multiple Gbs heaps).
 For interactive applications, a new algorithm is required with a smaller pause time.
@@
 
-First, we contacted multiple companies looking for use of the Cog VM in production with memory intensive applications (>1Gb heaps) and built benchmarks out of one of their deployed applications.
+First, we contacted multiple companies looking for production use of the Cog VM with memory intensive applications (>1Gb heaps) and built benchmarks out of one of the deployed applications.
 As an example, we will discuss in the paper in Section \ref{sec:mooseBench} the Moose benchmarks implemented with the help of Feenk\footnote{https://feenk.com/}.
@@
 
-\section{Building Benchmarks}
-\label{sec:bench}
+\section{Building Benchmarks} \label{sec:bench}
 \sk{isn't it weird to introduce the built benchmarks without explaining *why* they were chosen?
@@
 
-\subsection{Moose benchmark}
-\label{sec:mooseBench}
+\subsection{Moose benchmark} \label{sec:mooseBench}
 
@@
 
-\subsection{Configurable benchmark}
-\label{sec:confBench}
+\subsection{Configurable benchmark} \label{sec:confBench}
 
@@
 
-\section{Reference Implementations}
-\label{sec:ref}
+\section{Reference Implementations} \label{sec:ref}
 
@@
 We describe here the current Mark-Compact implementation, which has been in production for the past few years, and an implementation of Mark-Sweep we introduced as a reference earlier this year.
+Cog is written to allow new object representations to be written and used.
+Spur \cite{SpurMirandaBera} is the second object representation written for Cog.
+Spur avoids the cost of an explicit read barrier to follow forwarding pointers on most accesses by handling the forwarding check on the failure side of cached dynamic message send; any send to a forwarded object will fail, and forwarding pointers in the current frame are followed on the failure side of lookup.
+Hence in Spur all objects have room for a forwarding pointer.
+This is key to the implementation of the Mark-Compact algorithm.
 
-\subsection{Mark-Compact}
+\subsection{Mark-Compact} \label{sec:refmc}
 
@@
 
-\subsection{Mark-Sweep}
+\subsection{Mark-Sweep} \label{sec:refms}
 
@@
 
-\section{Reference Results}
-\label{sec:valid}
+\section{Reference Results} \label{sec:valid}
 
@@
 
-\caption{Moose benchmark results piled up (in ms).}
-\label{MooseRes1}
+\caption{Moose benchmark results piled up (in ms).} \label{MooseRes1}
 \end{figure*}
@@
 \end{tabular} 
-\caption{Moose benchmark results with standard errors.}
-\label{MooseRes2}
+\caption{Moose benchmark results with standard errors.} \label{MooseRes2}
 \end{figure*}

### 2018-06-20 paper.tex
add result description for large objects bench

--- 2018-06-20 paper.tex
@@
 
-\section{Introduction}
-\label{sec:intro}
+\section{Introduction} \label{sec:intro}
 
@@
 Then, new objects with a minimum size greater than the new free chunks sizes are allocated.
-\item \emph{LargeObjects:} Only large objects with a size greater than 0xFFFF \sk{put real size in Kb here} are allocated on the heap.
+\item \emph{LargeObjects:} Only large objects with a size greater than 65535 32-bits machines words are allocated on the heap. These objects are allocated directly in the old space.
 \item \emph{Data0bjects:} The majority of objects that will be allocated on the heap will be Data objects, which will only contains one-byte-sized integers of a specific size in this benchmark  
@@
 
-\subsection{Mark-Compact}
+\subsection{Mark-Compact} \label{sec:refmc}
 
@@
 \end{itemize}
-\paragraph{LargeObjects.}
-allocate objects larger than 0xffff. Compare two VMs with different number of segments.
-\begin{itemize}
-\item impact on memory size 
-\item impact on number of scavenges and number of full GC's
-\end{itemize}
+\paragraph{LargeObjects.} The LargeObjects benchmark allocates objects larger than 65535 32-bits machine words.
+The time spent in scavenges is therefore lower.
+Allocating large objects has an impact on the number operations of memory growing and shrinking, which can be lowered by tuning the associated parameters in the VM.
 \paragraph{Data0bjects.}

### 2018-06-20 paper.tex
fix previous merge
Remove Clement's obsolete note about vspace. Add the explanation of the three regimes of application/GC behaviour.

@@
 
-\cba{Maybe add vspace here not to have the title bottom of page}
+\section{Building Benchmarks} \label{sec:bench}
+Experience with applications deployed on Cog shows that memory intensive applications often exhibit three distinct regimes.
+As an application builds a new data structure the heap grows, and the garbage collector should be biassed to allow growth, rather than to favour performing garbage collection when the current size of the heap is reached.
+As an application processes the data structure it has built the heap is stable, at least in its size.
+If the application then releases significant storage and continues processing, the garbage collector should favour reclamation, and it is advantageous if it attempts to free up memory segments and return them to the OS.
+We have identified, therefore, growing, stable and shrinking regimes, and these are useful abstractions of real application behaviour, which are helpful in designing benchmarks and in providing tuning mechanisms to the system programmer.
 
-\section{Building Benchmarks} \label{sec:bench}
-\sk{isn't it weird to introduce the built benchmarks without explaining *why* they were chosen?
-Like, why growing, stable, shrinking is significant regarding garbage collection?
-same for my 3 benchs, why would I chose those 3 bench strategies?
-I know it is kind of explained in the result section but maybe it is a bit late}
-\cba{Sure, please explain}
 To build GC benchmarks, we took two approaches.

### 2018-06-20 paper.tex
Tweak designing benchmarks => designing meaningful benchmarks

@@
 If the application then releases significant storage and continues processing, the garbage collector should favour reclamation, and it is advantageous if it attempts to free up memory segments and return them to the OS.
-We have identified, therefore, growing, stable and shrinking regimes, and these are useful abstractions of real application behaviour, which are helpful in designing benchmarks and in providing tuning mechanisms to the system programmer.
+We have identified, therefore, growing, stable and shrinking regimes, and these are useful abstractions of real application behaviour, which are helpful in designing meaningful benchmarks and in providing tuning mechanisms to the system programmer.
 

### 2018-06-20 paper.tex
Use English spelling of bevaviour throughout.

@@
 Models currently used are encoded in up to 11Gb.
-This behavior was turned into three benchmarks, growing, accessing and shrinking the heap.
+This behaviour was turned into three benchmarks, growing, accessing and shrinking the heap.
 
@@
 We describe here the current Mark-Compact implementation, which has been in production for the past few years, and an implementation of Mark-Sweep we introduced as a reference earlier this year.
+
 Cog is written to allow new object representations to be written and used.
@@
 We can stress the garbage collector by running the configurable benchmarking tool with specific sets of options to test the memory startegy in edge cases.
-By increasing the amount of edge-cases benchmarks led on the garbage collector, we could infer a serie of tests to secure the garbage collector behavior in these situations.
+By increasing the amount of edge-cases benchmarks led on the garbage collector, we could infer a serie of tests to secure the garbage collector behaviour in these situations.
 

### 2018-06-20 paper.tex
Minor correction "64k 32-bits words" => "64k 32-bit words"

@@
 Then, new objects with a minimum size greater than the new free chunks sizes are allocated.
-\item \emph{LargeObjects:} Only large objects with a size greater than 65535 32-bits machines words are allocated on the heap. These objects are allocated directly in the old space.
+\item \emph{LargeObjects:} Only large objects with a size greater than 65535 32-bit machines words are allocated on the heap. These objects are allocated directly in the old space.
 \item \emph{Data0bjects:} The majority of objects that will be allocated on the heap will be Data objects, which will only contains one-byte-sized integers of a specific size in this benchmark  
@@
 \end{itemize}
-\paragraph{LargeObjects.} The LargeObjects benchmark allocates objects larger than 65535 32-bits machine words.
+\paragraph{LargeObjects.} The LargeObjects benchmark allocates objects larger than 65535 32-bit machine words.
 The time spent in scavenges is therefore lower.

### 2018-06-20 paper.tex
Missing period

@@
 The Linux VM was compiled using GCC version 5.4.0.
-The compiled VMs are 32 bit (x86) applications with 32-bit objects and pointers \footnote{Spur uses a common object header for both 32-bit and 64-bit versions, so 32-bit objects still have a 64-bit header and are all aligned on a 64-bit boundary.}
+The compiled VMs are 32 bit (x86) applications with 32-bit objects and pointers.
+\footnote{Spur uses a common object header for both 32-bit and 64-bit versions, so 32-bit objects still have a 64-bit header and are all aligned on a 64-bit boundary.}
 

### 2018-06-20 paper.tex
Consistently capitalise Eden.

@@
 It uses the fact that there is room for a forwarding pointer in all objects to store the eventual position of an object in the first field.
-It therefore first locates a large free chunk, or eden or a memory segment, to use as the savedFirstFieldsSpace, which it uses to store the first fields of objects that will be compacted, these fields being overwritten by each object's eventual position.
+It therefore first locates a large free chunk, or Eden or a memory segment, to use as the savedFirstFieldsSpace, which it uses to store the first fields of objects that will be compacted, these fields being overwritten by each object's eventual position.
 It then makes at least three passes through the heap.
@@
 \end{itemize}
-With a larger eden, more overall memory is used (in our case 90 extra Mb are used), but scavenge time is expected to be lower.
+With a larger Eden, more overall memory is used (in our case 90 extra Mb are used), but scavenge time is expected to be lower.
 Indeed, the scavenger is an implementation of Ungar's scavenger~\cite{Scavenger} and the scavenge time is proportional to the number of surviving objects.

### 2018-06-20 paper.tex
Fix Vm => VM
Use a concise cliche

@@
 With the compacting GC, memory goes back to its original size.
-With Sweep, multiple memory segments are left with a few live objects, preventing the Vm from returning those segments to the OS.
+With Sweep, multiple memory segments are left with a few live objects, preventing the VM from returning those segments to the OS.
 Hence, the memory decreases drastically, but not to the original size.
 Compaction time is much faster with Sweep, but as we have discussed, for less efficiency.
-Changing Eden size has not much, if any, impact here, since it's essentially a pure full GC benchmark.
+Changing Eden size has little if any impact here, since it's essentially a pure full GC benchmark.
 

### 2018-06-20 paper.tex
Better prhasing with a parenthesised remark.

@@
 \paragraph{ExpandPropertiesWithCache.} The ExpandProperties WithCache benchmark does not increase the overall memory size.
-Since the compacting GC has compacted the heap before starting the benchmark, there is enough room to allocate the large objects (large objects are allocated directly in old space) required by the benchmark and there is no need for a full GC during the benchmark.
+Since the compacting GC has compacted the heap before starting the benchmark, there is enough room to allocate the large objects required by the benchmark (large objects are allocated directly in old space) and there is no need for a full GC during the benchmark.
 However, with the sweep algorithm, there are multiple small free memory chunks which are not big enough to hold all the requested large objects, so a full GC is required.

### 2018-06-20 paper.tex
Configurable benchmark => Configurable Benchmark

@@
 
-\subsection{Configurable benchmark} \label{sec:confBench}
+\subsection{Configurable Benchmark} \label{sec:confBench}
 
@@
 
-\subsection{Configurable benchmark}
+\subsection{Configurable Benchmark}
 
-\paragraph{Set-up.}The Configurable benchmark evaluation was performed on an Asus Zenbook running Ubuntu 18.04 LTS, with a 2.2 Ghz processor Intel Core i5, and 8 Gb of 1600MHz DDR3 RAM.
+\paragraph{Set-up.}The Configurable Benchmark evaluation was performed on an Asus Zenbook running Ubuntu 18.04 LTS, with a 2.2 Ghz processor Intel Core i5, and 8 Gb of 1600MHz DDR3 RAM.
 The Linux VM was compiled using GCC version 5.4.0.

### 2018-06-20 paper.tex
Add brief mention of SpurSelectiveCompactor.  Now the reference to SpurSelectiveCompactor. in sec 3.3 "Snapshot discussion makes sense.

@@
 
+\subsection{Mark-SelectiveCompact}
+We added a new algorithm designed to reduce the pause time due to compaction.
+This algorithm, to be described fully elsewhere, chooses sparsely populated segments and compacts these to an empty segment to reduce fragmentation.
+It relies on Spur's implicit following of forwarding pointers to update references to moved objects without requiring a complete pass over the heap.
+
 \subsection{Snapshot discussion}

### 2018-06-20 paper.tex
New page before references.
The new algorithm is specifrically for SPur.  Say so.

@@
 \paragraph{Future work: incremental GC.}
-This investigation, including as it does a framework for compaction that includes a new store check, helps us in designing a fully incremental Mark-Sweep-Compact algorithm with very low pause times, suitable for interactive applications such as animation and gaming.
-
+This investigation, including as it does a framework for compaction that includes a new store check, helps us in designing a fully incremental Mark-Sweep-Compact algorithm for Spur with very low pause times, suitable for interactive applications such as animation and gaming.
+\newpage
 %% Bibliography

### 2018-06-20 paper.tex
Fix a couple of clumsy phrasings.

@@
 \end{itemize}
-With a larger Eden, more overall memory is used (in our case 90 extra Mb are used), but scavenge time is expected to be lower.
-Indeed, the scavenger is an implementation of Ungar's scavenger~\cite{Scavenger} and the scavenge time is proportional to the number of surviving objects.
+With a larger Eden, more overall memory is used (in our case an extra 90 Mb), but scavenge time is expected to be lower.
+The Spur scavenger is an extended implementation of Ungar's classic generation scavenger~\cite{Scavenger} and the scavenge time is proportional to the number of surviving objects.
 %To evaluate the performance, we use:
@@
 We believe this is due to worse locality of objects.
-With a larger Eden, the scavenge time is much smaller and the execution time is a bit faster due to improved object locality.
+With a larger Eden,  scavenge time is much smaller and the execution time is a bit faster due to improved object locality.
 

### 2018-06-20 paper.tex
update parts on Data Objects

@@
 
-First, we contacted multiple companies looking for production use of the Cog VM with memory intensive applications (>1Gb heaps) and built benchmarks out of one of the deployed applications.
+First, we contacted multiple companies looking for production use of the Cog VM with memory intensive applications ($>$1Gb heaps) and built benchmarks out of one of the deployed applications.
 As an example, we will discuss in the paper in Section \ref{sec:mooseBench} the Moose benchmarks implemented with the help of Feenk\footnote{https://feenk.com/}.
@@
 \item \emph{LargeObjects:} Only large objects with a size greater than 65535 32-bit machines words are allocated on the heap. These objects are allocated directly in the old space.
-\item \emph{Data0bjects:} The majority of objects that will be allocated on the heap will be Data objects, which will only contains one-byte-sized integers of a specific size in this benchmark  
+\item \emph{Data0bjects:} The majority of the objects that will be allocated on the heap will be Data objects, which will only contain one-byte-sized integers in this benchmark.
+Data Objects can contain integers, float or double: the garbage collector will not go through their slots in the marking phase.
 \end{itemize}
@@
 Allocating large objects has an impact on the number operations of memory growing and shrinking, which can be lowered by tuning the associated parameters in the VM.
-\paragraph{Data0bjects.}
- allocate a lot of data objects (90\% of all allocated objects during the bench). Compare for Mark-Sweep and Mark-Compact
-\begin{itemize}
-\item impact on memory size
-\item impact on overall time (full gc, compaction time)
-\end{itemize}
+\paragraph{Data0bjects.} The DataObjects benchmark allocates a large amount of data objects on the heap (90\% of all allocated objects during the benchmark).
+The time spent in full GC's and the compaction time are lower in comparison with a case with 40\% of Data Objects.
+This can be explained because less time is spent in the marking and pointer update phases.
 

### 2018-06-20 paper.tex
Minor edits

@@
 Then, new objects with a minimum size greater than the new free chunks sizes are allocated.
-\item \emph{LargeObjects:} Only large objects with a size greater than 65535 32-bit machines words are allocated on the heap. These objects are allocated directly in the old space.
+\item \emph{LargeObjects:} Only large objects with a size greater than 65535 32-bit machine words are allocated on the heap. These objects are allocated directly in the old space.
 \item \emph{Data0bjects:} The majority of the objects that will be allocated on the heap will be Data objects, which will only contain one-byte-sized integers in this benchmark.
@@
 The Linux VM was compiled using GCC version 5.4.0.
-The compiled VMs are 32 bit (x86) applications with 32-bit objects and pointers.
+The compiled VMs are 32-bit (x86) applications with 32-bit objects and pointers.
 \footnote{Spur uses a common object header for both 32-bit and 64-bit versions, so 32-bit objects still have a 64-bit header and are all aligned on a 64-bit boundary.}

### 2018-06-20 paper.tex
Rephrase number operations of memory...

@@
 The time spent in scavenges is therefore lower.
-Allocating large objects has an impact on the number operations of memory growing and shrinking, which can be lowered by tuning the associated parameters in the VM.
+Allocating large objects has an impact on the number of memory growing and shrinking operations, which can be altered by tuning the associated parameters in the VM.
 \paragraph{Data0bjects.} The DataObjects benchmark allocates a large amount of data objects on the heap (90\% of all allocated objects during the benchmark).

### 2018-06-20 paper.tex
time spent in scavenges => time spent in scavenging

@@
 Total full GC time is the total time spent in full GCs, including  compaction time.
-Scavenge time is the time spent in scavenges, \emph{i.e.,} garbage collection of young objects only.
+Scavenge time is the time spent in scavenging, \emph{i.e.,} garbage collection of young objects only.
 Total execution time is the total execution time, including scavenge and full GC time.
@@
 allocate n objects of size x, release half of them, and reallocate n/2 objects of size y>x.
-Compare the performance (time spent in full GC, scavenges, compacting, and execution time) for both Mark-Sweep and Mark-Compact
+Compare the performance (time spent in full GC, scavenging, compacting, and execution time) for both Mark-Sweep and Mark-Compact
 \begin{itemize}
@@
 \paragraph{LargeObjects.} The LargeObjects benchmark allocates objects larger than 65535 32-bit machine words.
-The time spent in scavenges is therefore lower.
+The time spent in scavenging is therefore lower.
 Allocating large objects has an impact on the number of memory growing and shrinking operations, which can be altered by tuning the associated parameters in the VM.

### 2018-06-20 paper.tex
add last figures

--- 2018-06-20 paper.tex
@@
 These options can be associated with different heap states: growing, stable or shrinking, like in our other approach.
-We ran three benchmark cases with this tool: fragmented heap, large objects in heap and majority of data objects in heap.
+We ran two benchmark cases with this tool: large objects in heap and majority of data objects in heap.
 This implementation and the use-cases are detailed in Section \ref{sec:confBench}
@@
 
-We built three benchmarks out of the potential configurations:
+We built two benchmarks out of the potential configurations:
 \begin{itemize}
-\item \emph{FragmentedHeap:} First, objects are allocated on the heap, and some of them are released, leading to potential memory fragmentation.
+%\item \emph{FragmentedHeap:} First, objects are allocated on the heap, and some of them are released, leading to potential memory fragmentation.
 Then, new objects with a minimum size greater than the new free chunks sizes are allocated.
@@
 \item \emph{Data0bjects:} The majority of the objects that will be allocated on the heap will be Data objects, which will only contain one-byte-sized integers in this benchmark.
-Data Objects can contain integers, float or double: the garbage collector will not go through their slots in the marking phase.
+Data Objects are raw data objects, which can be used, for example, to store characters, restricted-range integers, floats, etc: the garbage collector will not go through their slots in the marking phase.
 \end{itemize}
@@
 
+\begin{figure*}[b!]
+	\centering
+
+    \begin{subfigure}[b]{.48\textwidth}
+	\includegraphics[width=\linewidth]{figures/big} 
+	\caption{LargeObjects\vspace{0.2cm}}
+   	\end{subfigure}\hspace{0.03\textwidth}% 
+   	\begin{subfigure}[b]{.48\textwidth}
+	\includegraphics[width=\linewidth]{figures/data} 
+	\caption{DataObjects\vspace{0.2cm}}
+   	\end{subfigure}
+
+\caption{Configurable benchmark results piled up (in ms).} \label{ConfigRes1}
+\end{figure*}
+
+\begin{figure*}[b!]
+\begin{tabular}{|c|c|l|l|l|l|}
+   \hline
+  & & Total   & Scavenge  & Total Full  & Compaction   \\
+ Benchmark & &  Exec. time &  time & GC time &  time \\
+  & &  (ms) &  (ms) & (ms) &  (ms)\\
+   \hline
+   \multirow{4}{*}{LargeObjects} & Small 	& 43110 $\pm$ 338 &	3223 $\pm$ 76 	& 541 $\pm$ 5 	&298 $\pm$ 2.8  \\
+    				    & Small-NewMem 	& 42886 $\pm$ 341 &	3204 $\pm$ 23 	& 605 $\pm$ 43 	&348 $\pm$ 8  \\
+    				    & Large 	& 1514 $\pm$ 15 &	155 $\pm$ 4 	& 414 $\pm$ 1 	&223 $\pm$ 4  \\
+    				    & Large-NewMem 		& 1506 $\pm$ 20 &	163 $\pm$ 8 	& 408 $\pm$ 6 	&219 $\pm$ 6  \\
+   \hline
+   \multirow{4}{*}{DataObjects} 	& Compact-Less 		& 55757 $\pm$ 1998 &	6465 $\pm$ 122 	& 403 $\pm$ 36 	&232 $\pm$ 20.55  \\
+    				    			& Compact 	& 12738 $\pm$ 13090 &	569 $\pm$ 10 	& 168 $\pm$ 27 	&98 $\pm$ 15  \\
+    				    & Sweep 	& 13785 $\pm$ 269 &	604 $\pm$ 9 	& 115 $\pm$ 2 	&27 $\pm$ 1 \\
+    				    & Sweep-Less 		& 59398 $\pm$ 592 &	6679 $\pm$ 34 	& 226 $\pm$ 3 	&56 $\pm$ 1 \\
+
+   \hline
+\end{tabular} 
+\caption{Configurable benchmark results with standard errors.} \label{ConfigRes2}
+\end{figure*}
+
 \paragraph{Set-up.}The Configurable Benchmark evaluation was performed on an Asus Zenbook running Ubuntu 18.04 LTS, with a 2.2 Ghz processor Intel Core i5, and 8 Gb of 1600MHz DDR3 RAM.
@@
 
-\paragraph{FragmentedHeap.} 
-allocate n objects of size x, release half of them, and reallocate n/2 objects of size y>x.
-Compare the performance (time spent in full GC, scavenges, compacting, and execution time) for both Mark-Sweep and Mark-Compact
-\begin{itemize}
-\item impact on memory size and numbers of segments
-\item impact on execution time 
-\end{itemize}
-\paragraph{LargeObjects.} The LargeObjects benchmark allocates objects larger than 65535 32-bit machine words.
-The time spent in scavenges is therefore lower.
+%\paragraph{FragmentedHeap.} The fragmented heap benchmark allocates objects of 300 slots maximum, releases half of them, and re-allocates half of the amount of objects with a minimum size of 600 slots.
+
+\paragraph{LargeObjects.} The LargeObjects benchmark allocates objects larger than 65535 32-bits machine words.
+The time spent in scavenges is therefore lower than with smaller objects (that are not allocated in the old space at first).
 Allocating large objects has an impact on the number operations of memory growing and shrinking, which can be lowered by tuning the associated parameters in the VM.
+In this example, the New-Mem cases have memory segments four times bigger, and four empty segments are available instead of one.
 \paragraph{Data0bjects.} The DataObjects benchmark allocates a large amount of data objects on the heap (90\% of all allocated objects during the benchmark).
@@
 The nursery (new space) survival rate is also taken into account, as well as the heap structure through object lifetime behavior.
-These criteria were identified as significant in assessing the performance of memory management \eem{by whom?}.
-
-%The part dedicated to memory management focuses on three metrics: allocation demographics, lifetime and live objects.
-The benchmarks used shows a rich and diverse object lifetime behaviors.
---> dacapo allocates more objets, 3 time the live size of SPEC.
-%Heap volume, Heap objects, Mean object size, nursery survival rate.
-%
-%measure object size demographics (size demographics for each benchmark)
-%--> live and allocation time serie analysis
-%plot heap composition (allocations and pointer mutations)
-%--> richer lifetime behavior
 

### 2018-06-20 paper.tex

@@
 \begin{itemize}
-%\item \emph{FragmentedHeap:} First, objects are allocated on the heap, and some of them are released, leading to potential memory fragmentation.
-Then, new objects with a minimum size greater than the new free chunks sizes are allocated.
 \item \emph{LargeObjects:} Only large objects with a size greater than 65535 32-bit machine words are allocated on the heap. These objects are allocated directly in the old space.
@@
 
-\subsection{Configurable Benchmark}
-
 \begin{figure*}[b!]
@@
 
+\subsection{Configurable Benchmark}
+
 \paragraph{Set-up.}The Configurable Benchmark evaluation was performed on an Asus Zenbook running Ubuntu 18.04 LTS, with a 2.2 Ghz processor Intel Core i5, and 8 Gb of 1600MHz DDR3 RAM.
@@
 
-%\paragraph{FragmentedHeap.} The fragmented heap benchmark allocates objects of 300 slots maximum, releases half of them, and re-allocates half of the amount of objects with a minimum size of 600 slots.
-
-\paragraph{LargeObjects.} The LargeObjects benchmark allocates objects larger than 65535 32-bits machine words.
-The time spent in scavenges is therefore lower than with smaller objects (that are not allocated in the old space at first).
-Allocating large objects has an impact on the number operations of memory growing and shrinking, which can be lowered by tuning the associated parameters in the VM.
-In this example, the New-Mem cases have memory segments four times bigger, and four empty segments are available instead of one.
+\paragraph{LargeObjects.} The LargeObjects benchmark allocates objects larger than 65535 32-bit machine words.
+The time spent in scavenging is therefore lower.
+Allocating large objects has an impact on the number of memory growing and shrinking operations, which can be altered by tuning the associated parameters in the VM.
 \paragraph{Data0bjects.} The DataObjects benchmark allocates a large amount of data objects on the heap (90\% of all allocated objects during the benchmark).

### 2018-08-20 paper.tex
take reviews into account - graph titles, add quote for 2 finger algorithm, fix typos

@@
   \department{Masters student}              %% \department is recommended
-  \institution{Universit\'e de Lille 1}            %% \institution is required
+  \institution{Universit\'e Lille 1}            %% \institution is required
  % \streetaddress{Street1 Address1}
@@
 Part of the business of Feenk consists in analyzing large software systems.
-To do so, the application parses a mse file into a model, performs some analysis, and then releases the model and generates analysis results.
+To do so, the application parses a mse file into a model (mse is the default file format supported by Moose), performs some analysis, and then releases the model and generates analysis results.
 For business, the application analyzes closed-sourced application of companies.
@@
 It uses the fact that there is room for a forwarding pointer in all objects to store the eventual position of an object in the first field.
-It therefore first locates a large free chunk, or Eden or a memory segment, to use as the savedFirstFieldsSpace, which it uses to store the first fields of objects that will be compacted, these fields being overwritten by each object's eventual position.
+It therefore first locates a large free chunk, or the Eden space or a memory segment, to use as the savedFirstFieldsSpace, which it uses to store the first fields of objects that will be compacted, these fields being overwritten by each object's eventual position.
 It then makes at least three passes through the heap.
@@
 
-Each pass uses a three finger algorithm, a simple extension of the classic two finger algorithm with an extra middle finger used to refer to the lowest pinned object, if any, between the to and from fingers.
+Each pass uses a three finger algorithm, a simple extension of the classic two finger algorithm \cite{Cheney} with an extra middle finger used to refer to the lowest pinned object, if any, between the to and from fingers.
 Objects are moved down, starting at the first free object or chunk, provided that they fit below the lowest pinned object above the to finger.
@@
 	\includegraphics[width=\linewidth]{figures/load} 
-	\caption{Load\vspace{0.2cm}}
+	\caption{LoadFromMSE\vspace{0.2cm}}
    	\end{subfigure}\hspace{0.03\textwidth}% 
@@
    \hline
-   \multirow{4}{*}{Load} & Sweep 	& 159821 $\pm$ 2218 &	73599 $\pm$ 1281 	& 10863 $\pm$ 726 	&1398 $\pm$ 91 	& 193 $\pm$ 0 	& 959 $\pm$ 9.69 \\
+   \multirow{4}{*}{LoadFromMSE} & Sweep 	& 159821 $\pm$ 2218 &	73599 $\pm$ 1281 	& 10863 $\pm$ 726 	&1398 $\pm$ 91 	& 193 $\pm$ 0 	& 959 $\pm$ 9.69 \\
     				    & Compact 	& 162305 $\pm$ 2790 &	72150 $\pm$ 782 	& 14091 $\pm$ 972 	&4661 $\pm$ 289 	& 193 $\pm$ 0 	& 909 $\pm$ 9.69 \\
@@
 
-\paragraph{Load.} The Load benchmark grows the memory by around 700Mb.
+\paragraph{LoadFromMSE.} The LoadFromMSE benchmark grows the memory by around 700Mb.
 Compaction time is smaller for the sweep algorithm, making the benchmark a little bit faster in the Mark-Sweep, which is to be expected since Sweep does not need to waste time updating pointers because objects are not moved in memory.

### 2018-08-20 paper.tex
add dexcription of the 4 configurations for largeObjects bench

@@
 We evaluated the configurable benchmarks using the reference Mark-Compact and Mark-Sweep algorithms and by generating specific heap configurations. 
-The results are shown in Figure \ref{ConfigRes1} and \ref{ConfigRes2}
+The results are shown in Figure \ref{ConfigRes1} and \ref{ConfigRes2}.
 
-\paragraph{LargeObjects.} The LargeObjects benchmark allocates objects larger than 65535 32-bit machine words.
-The time spent in scavenging is therefore lower.
+\paragraph{LargeObjects.} The LargeObjects benchmark compare the time spent in garbage collection according to the size of allocated objects.
+ We distinguish four configurations:
+\begin{itemize}
+\item \emph{Small}: VM with standard GC parameters and the Mark-Compact algorithm. Objects smaller than 65535 32-bit machine words are allocated.
+\item \emph{Small-NewMem}: VM with standard GC parameters, except segments size which is 4 times larger, the number of empty segments which is multiplied by 4, and the Mark-Compact algorithm.
+Objects smaller than 65535 32-bit machine words are allocated.
+\item \emph{Large}: VM with standard GC parameters and the Mark-Compact algorithm. Objects larger than 65535 32-bit machine words are allocated.
+\item \emph{Large-NewMem}: VM with standard GC parameters, except segments size which is 4 times larger, the number of empty segments which is multiplied by 4, and the Mark-Compact algorithm.
+Objects larger than 65535 32-bit machine words are allocated.
+\end{itemize}
+The time spent in scavenging is lower for the larger objects.
 Allocating large objects has an impact on the number of memory growing and shrinking operations, which can be altered by tuning the associated parameters in the VM.

### 2018-08-20 paper.tex
add dexcription of the 4 configurations for dataobjects bench

@@
 
-\paragraph{LargeObjects.} The LargeObjects benchmark compare the time spent in garbage collection according to the size of allocated objects.
- We distinguish four configurations:
+\paragraph{LargeObjects.} The LargeObjects benchmark compares the time spent in garbage collection according to the size of allocated objects. We distinguish four configurations:
 \begin{itemize}
@@
 Allocating large objects has an impact on the number of memory growing and shrinking operations, which can be altered by tuning the associated parameters in the VM.
-\paragraph{Data0bjects.} The DataObjects benchmark allocates a large amount of data objects on the heap (90\% of all allocated objects during the benchmark).
+\paragraph{Data0bjects.} The DataObjects benchmark compares the time spent in garbage collection according to the amount of allocated data objects.
+We distinguish four configurations:
+\begin{itemize}
+\item \emph{Compact-Less}: VM with standard GC parameters and the Mark-Compact algorithm. 40\% of allocated objects are data objects.
+\item \emph{Compact}: VM with standard GC parameters and the Mark-Compact algorithm. 90\% of allocated objects are data objects.
+\item \emph{Sweep-Less}: VM with standard GC parameters and the Mark-Sweep algorithm. 40\% of allocated objects are data objects.
+\item \emph{Sweep}: VM with standard GC parameters and the Mark-Sweep algorithm. 
+90\% of allocated objects are data objects.
+\end{itemize}
  The time spent in full GC's and the compaction time are lower in comparison with a case with 40\% of Data Objects.

### 2018-08-21 paper.tex
replace blurry histograms
add acknowledgments
fix typo

@@
 
-    \begin{subfigure}[b]{.48\textwidth}
-	\includegraphics[width=\linewidth]{figures/big} 
+    \begin{subfigure}[b]{.499\textwidth}
+	\includegraphics[width=\linewidth]{figures/large} 
 	\caption{LargeObjects\vspace{0.2cm}}
-   	\end{subfigure}\hspace{0.03\textwidth}% 
-   	\begin{subfigure}[b]{.48\textwidth}
+   	\end{subfigure}\hspace{0.00000001\textwidth}% 
+   	\begin{subfigure}[b]{.499\textwidth}
 	\includegraphics[width=\linewidth]{figures/data} 
@@
 \item \emph{Small}: VM with standard GC parameters and the Mark-Compact algorithm. Objects smaller than 65535 32-bit machine words are allocated.
-\item \emph{Small-NewMem}: VM with standard GC parameters, except segments size which is 4 times larger, the number of empty segments which is multiplied by 4, and the Mark-Compact algorithm.
+\item \emph{Small-NewMem}: VM with standard GC parameters, except segment size which is 4 times larger, number of empty segments which is multiplied by 4, and the Mark-Compact algorithm.
 Objects smaller than 65535 32-bit machine words are allocated.
 \item \emph{Large}: VM with standard GC parameters and the Mark-Compact algorithm. Objects larger than 65535 32-bit machine words are allocated.
-\item \emph{Large-NewMem}: VM with standard GC parameters, except segments size which is 4 times larger, the number of empty segments which is multiplied by 4, and the Mark-Compact algorithm.
+\item \emph{Large-NewMem}: VM with standard GC parameters, except segment size which is 4 times larger, number of empty segments which is multiplied by 4, and the Mark-Compact algorithm.
 Objects larger than 65535 32-bit machine words are allocated.
@@
 This investigation, including as it does a framework for compaction that includes a new store check, helps us in designing a fully incremental Mark-Sweep-Compact algorithm for Spur with very low pause times, suitable for interactive applications such as animation and gaming.
-\newpage
+
+\section*{Acknowledgments}
+We thank Tudor Girba for providing the files needed for the Moose benchmark.
+
 %% Bibliography
